### 2024-11-26 paper/paper.qmd
update data section

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Xizi Sun
thanks: "Code and data are available at: https://github.com/LilianS77/US_Marriage."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(ggplot2)
library(dplyr)
library(here)
library(arrow)
library(scales)
library(knitr)
```

```{r}
#| echo: false
#| warning: false
#| message: false

data <- read_parquet(here("data", "01-analysis_data", "analysis_data.parquet"))

```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Data Tool

The dataset was analyzed using R [@citeR] and utilized several R packages for data manipulation and visualization, including ggplot2 [@citeGGplot2], dplyr [@citeDplyr], and here [@citeHere]. The data was processed using the Apache Arrow package [@citeArrow] for efficient file handling and visualized with scales [@citeScales] for enhanced graphics. Reproducibility was ensured using knitr [@citeKnitr]. Data was extracted from IPUMS USA [@ipumsusa]. Guidance on storytelling with data was drawn from Telling Stories with Data [@tellingstories].

## Data Source And Measurement

The data used for this study was sourced from **IPUMS USA (Integrated Public Use Microdata Series)** [@ipumsusa], a comprehensive repository providing access to harmonized census and survey data from the United States. IPUMS USA is renowned for its meticulous process of standardizing variables across datasets, ensuring compatibility over time, and providing extensive metadata to support research. This harmonization process enables researchers to conduct longitudinal and cross-sectional studies on social, demographic, and economic trends.

The dataset consists of **individual-level microdata**, where each record represents a single person, numerically coded for all relevant characteristics. These records are organized into households, allowing for the study of individual behaviors and characteristics in the context of their family or co-residential settings. Unlike compiled statistics or pre-aggregated tables, this microdata structure provides researchers with unparalleled flexibility in exploring relationships between variables.

To address the diversity of record layouts, coding schemes, and documentation across the historical scope of the dataset, IPUMS implements a rigorous **harmonization process**. Variables are assigned **uniform codes**, ensuring consistency across census years (1850–2010) and the American Community Surveys (ACS) (2000–present). This standardization simplifies the analysis of long-term trends and facilitates comparisons across time and space.

## Variable Selection

Using the IPUMS data extraction system, this study selected a focused subset of variables to examine the determinants of **non-marriage** in the United States. These include key demographic and socioeconomic characteristics, which are numerically coded for statistical analysis.

The table [@tbl-cleaned-data] shows the variables after data cleaning,

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of cleaned data
#| echo: false
#| warning: false
#| message: false

# Display the first 6 rows of the dataset with column names
head(data[, c("marital_status", "age", "gender", "Race", "Income", "education_level")], 6) |>
  kable(
    col.names = c("Marital Status", "Age", "Gender", "Race", "Income", "Education Level"),
    booktabs = TRUE
  )
```

### Outcome variables
The primary outcome variable for this study is **Marital Status**, which categorizes individuals based on their marital state. The proportion of marital status categories is displayed in figure @fig-proportion-marriage. This variable allows for a comparison between individuals who have never married (`Not_Married`) and those who have (`Married`). For this study,
**Not_Married**: Includes individuals who have never been married.
**Married**: Includes individuals who are married, as well as those who are divorced, widowed, or separated.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-proportion-marriage
#| fig-cap: Proportion of Marital Status Categories
#| 
# Calculate proportions
marital_status_counts <- table(data$marital_status)
marital_status_df <- as.data.frame(marital_status_counts)
colnames(marital_status_df) <- c("Marital_Status", "Count")
marital_status_df$Proportion <- marital_status_df$Count / sum(marital_status_df$Count)

# Plot
ggplot(marital_status_df, aes(x = Marital_Status, y = Proportion, fill = Marital_Status)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Proportion of Marital Status Categories",
    x = "Marital Status",
    y = "Proportion"
  ) +
  theme_minimal()
```

### Predictor variables

This section focuses on the predictor variables included in the study. The distribution of predictor variables is displayed in Figure @fig-predictor-variables. These variables capture demographic, socioeconomic, and personal characteristics, providing a comprehensive framework for analyzing factors associated with marital status. Below are the key predictor variables:

1. **Age**: A continuous variable representing the respondent's age in years.
2. **Gender**: A categorical variable indicating whether the respondent is male or female. Gender differences often play a role in marital patterns.
3. **Race**: A categorical variable categorized into White, Black, Asian, American Indian, and Other racial groups. This variable examines potential racial disparities in marital behavior.
4. **Education Level**: An ordinal variable indicating the highest level of education attained by the respondent. It is grouped into five categories: Below High School, High School, Some College, Bachelor’s Degree, and Above Bachelor.
5. **Income**: A continuous variable measuring the respondent's annual income in dollars. Income reflects economic resources and may be associated with marital stability and decisions.

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: fig-predictor-variables
#| fig-subcap: ["Age Distribution", "Gender Distribution", "Race Distribution", "Education Level Distribution", "Income Distribution"]
#| label: fig-predictor-variables
#| layout-ncol: 2

# Function to create a bar plot or histogram
create_plot <- function(data, x_var, x_label, is_continuous = FALSE, binwidth = NULL) {
  if (is_continuous) {
    # Create a histogram for continuous variables
    p <- ggplot(data, aes_string(x = x_var)) +
      geom_histogram(binwidth = binwidth, fill = "skyblue", color = "black") +
      labs(x = x_label, y = "Count") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  } else {
    # Create a bar chart for categorical variables
    p <- ggplot(data, aes_string(x = x_var)) +
      geom_bar(fill = "skyblue", color = "black") +
      labs(x = x_label, y = "Count") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  }
  return(p)
}

# Create plots for the predictor variables
create_plot(data, "age", "Age", is_continuous = TRUE, binwidth = 5)
create_plot(data, "gender", "Gender")
create_plot(data, "Race", "Race")
create_plot(data, "education_level", "Education Level")
create_plot(data, "Income", "Income", is_continuous = TRUE, binwidth = 10000)

```

## Data Selection

To understand the phenomenon of non-marriage in the United States, I selected the IPUMS USA dataset over alternatives such as IPUMS International. While IPUMS International provides harmonized census data from 104 countries with over 1 billion person records, its vast scope makes it less targeted for this study. My focus is specifically on the U.S. population, and IPUMS USA offers high-precision data from American Community Surveys (ACS) and federal censuses, making it a more suitable choice for studying societal patterns specific to the United States.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-Income-Unmarried
#| fig-cap: Income Distribution for Unmarried Individuals

# Filter for unmarried individuals
unmarried_data <- data %>% filter(marital_status == "Not_Married")

# Create the income distribution plot without scientific notation
ggplot(unmarried_data, aes(x = Income)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black", alpha = 0.7) +
  geom_density(color = "blue", size = 1) +
  ggtitle("Income Distribution for Unmarried Individuals") +
  xlab("Income") +
  ylab("Density") +
  scale_x_continuous(labels = comma) + # Format x-axis labels as regular numbers
  theme_minimal()
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-Age-Unmarried
#| fig-cap: Age Distribution for Unmarried Individua

# Filter for unmarried individuals
unmarried_data <- data %>% filter(marital_status == "Not_Married")

# Create the age distribution plot
ggplot(unmarried_data, aes(x = age)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "salmon", color = "black", alpha = 0.7) +
  geom_density(color = "red", size = 1) +
  ggtitle("Age Distribution for Unmarried Individuals") +
  xlab("Age") +
  ylab("Density") +
  theme_minimal()
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix

## How to Extract 2022 ACS Data from IPUMS

To obtain data from IPUMS, we start by navigating to the IPUMS USA section and clicking on Get Data. Next, we go to the Select Sample section, where we uncheck the "Default sample from each year" option and instead select 2023 ACS. After selecting our sample, we proceed to add variables of interest. For individual-level data, we might add variables from the Person section. For example, under Demographic, we could include variables like AGE, and under Person, we could add SEX, RACE, INCTOT (total personal income) and EDUC (education attainment). Once our variables are selected, we click View Cart, then proceed by clicking Create Data Extract. At this point, we review our selections, change the Data Format to CSV, and submit our extract for processing. Then we saved it locally as usa_00001.csv.

\newpage

# References

### 2024-11-26 paper/paper.qmd
update model draft

@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+For this analysis, I employed a Logistic Regression Model to assess the likelihood of an individual not being married (outcome variable) based on several demographic and socioeconomic predictors (predictor variables).
+This model was chosen due to the binary nature of the outcome variable, which distinguishes between individuals who are "Not Married" versus those who are married (including divorced, widowed, or separated)..
 
@@
 
-## Model set-up
+## Model Setup
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+### Objective
+The primary objective of the model is to analyze and predict the factors associated with individuals' marital status, focusing on identifying key predictors for individuals who have never been married (`Not_Married`).
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+### Model Type
+A **logistic regression model** is chosen for this analysis, as the outcome variable (`marital_status`) is binary (e.g., `Not_Married` vs. `Married`).
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+### Outcome Variable
+- **`marital_status`**: Encoded as:
+  - `1` for `Not_Married`
+  - `0` for `Married`
+
+### Predictor Variables
+The predictors include:
+1. **Age (`age`)**: A continuous variable representing the individual’s age.
+2. **Gender (`gender`)**: A categorical variable with levels `Male` and `Female`.
+3. **Race (`Race`)**: A categorical variable with multiple racial categories (e.g., `White`, `Black`, etc.).
+4. **Income (`Income`)**: A continuous variable representing the individual’s annual income.
+5. **Education Level (`education_level`)**: An ordinal variable with levels `Below_High_School`, `High_School`, `Some_College`, `Bachelor`, and `Above_Bachelor`.
+
+### Model Assumptions
+1. **Linearity**: The log odds of the outcome are linearly related to the predictors.
+2. **Independence**: Observations are independent of each other.
+3. **No Multicollinearity**: Predictors are not highly correlated.
+4. **No Outliers or Influential Points**: Checked through residual analysis.
+
+### Mathematical Representation
+
+$$
+\log\left(\frac{P(\text{Not\_Married})}{1 - P(\text{Not\_Married})}\right) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{gender} + \beta_3 \cdot \text{Race} + \beta_4 \cdot \text{Income} + \beta_5 \cdot \text{education\_level}
+$$
+
+Where:
+- \( P(\text{Not\_Married}) \) is the probability of being `Not_Married`.
+
+### Model Implementation Steps
+1. **Data Preprocessing**
+   - Encode categorical variables (e.g., `gender`, `Race`, `education_level`) using one-hot encoding or dummy variables.
+   - Normalize continuous variables like `age` and `Income` for better model performance.
+   - Split the data into training and testing sets (e.g., 80% train, 20% test).
+
+2. **Model Building**
+   - Use the `glm` function in R for logistic regression modeling.
+   - Define the formula as:
+     ```r
+     glm(marital_status ~ age + gender + Race + Income + education_level, 
+         data = training_data, 
+         family = "binomial")
+     ```
+
+3. **Validation and Performance**
+   - Evaluate the model using metrics such as:
+     - **Accuracy**: Proportion of correctly predicted observations.
+     - **ROC Curve and AUC**: To assess the tradeoff between sensitivity and specificity.
+     - **Confusion Matrix**: To evaluate true positives, true negatives, false positives, and false negatives.
+
+4. **Interpretation**
+   - Analyze the coefficients (\( \beta \)) to determine the relationship between predictors and the likelihood of being `Not_Married`.
+
+### Potential Limitations
+- Overfitting with too many predictors.
+- Missing data or imbalanced classes in the `marital_status` variable.
 
 ### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
 
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
 # Discussion

### 2024-11-26 paper/paper.qmd
update model set-up

--- 2024-11-26 paper/paper.qmd
@@
 
-## Model set-up
+## Model Setup
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+### Objective
+The primary objective of the model is to analyze and predict the factors associated with individuals' marital status, focusing on identifying key predictors for individuals who have never been married (`Not_Married`).
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+## Model Setup
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+The logistic regression model used in this study is:
+
+$$
+\log \left( \frac{\hat{p}}{1 - \hat{p}} \right) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{gender} + \beta_3 \cdot \text{Race} + \beta_4 \cdot \text{Income} + \beta_5 \cdot \text{education\_level}
+$$
+
+The priors for the coefficients are defined as:
+
+$$
+\begin{aligned}
+\beta_0 &\sim \text{Normal}(0, 2.5), \\
+\beta_1 &\sim \text{Normal}(0, 2.5), \\
+\beta_2 &\sim \text{Normal}(0, 2.5), \\
+\beta_3 &\sim \text{Normal}(0, 2.5), \\
+\beta_4 &\sim \text{Normal}(0, 2.5), \\
+\beta_5 &\sim \text{Normal}(0, 2.5). \\
+\end{aligned}
+$$
+
+Where:
+
+- $\hat{p}$: Represents the probability that an individual is classified as `Not_Married`.
+- $\beta_0$: The intercept term, representing the log-odds of being `Not_Married` when all predictors are zero.
+- $\beta_1$: The coefficient for the individual's age.
+- $\beta_2$: The coefficient for gender.
+- $\beta_3$: The coefficient for the race category.
+- $\beta_4$: The coefficient for annual income.
+- $\beta_5$: The coefficient for the education level.
+
+### Priors Explanation
+In this model, normal priors with a mean of 0 and a standard deviation of 2.5 are assigned to all coefficients and the intercept.
+This choice reflects neutrality, implying no prior expectation of direction or magnitude for the coefficients.
+
+The standard deviation of 2.5 reflects moderate uncertainty in prior beliefs, allowing the model to incorporate reasonable variability in predictor effects while avoiding overly restrictive priors.
 
 ### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
 

### 2024-11-26 paper/paper.qmd
update model section

@@
 For this analysis, I employed a Logistic Regression Model to assess the likelihood of an individual not being married (outcome variable) based on several demographic and socioeconomic predictors (predictor variables).
-This model was chosen due to the binary nature of the outcome variable, which distinguishes between individuals who are "Not Married" versus those who are married (including divorced, widowed, or separated)..
+This model was chosen due to the binary nature of the outcome variable, which distinguishes between individuals who are "Not Married" versus those who are married (including divorced, widowed, or separated).
 
@@
 The primary objective of the model is to analyze and predict the factors associated with individuals' marital status, focusing on identifying key predictors for individuals who have never been married (`Not_Married`).
-
-## Model Setup
 
@@
 ### Priors Explanation
+
 In this model, normal priors with a mean of 0 and a standard deviation of 2.5 are assigned to all coefficients and the intercept.
@@
 
-### Model justification
+## Model justification
+
+The logistic regression model was chosen for this study due to its suitability for predicting binary outcomes, in this case, whether an individual is categorized as "Not Married" or not.
+Logistic regression is widely used in social science research to understand relationships between a binary dependent variable and multiple independent variables, making it an ideal choice for analyzing marital status.
+This model allows for a straightforward interpretation of the effects of predictors, such as age, income, education, and gender, through odds ratios, providing insight into the likelihood of being "Not Married." The dataset contains a mix of continuous and categorical variables, and logistic regression can handle this diversity effectively.
+Additionally, its theoretical simplicity ensures that the analysis remains interpretable while maintaining robustness.
+The choice of this model aligns with the study’s goal of identifying significant factors associated with non-marriage in the U.S., providing actionable insights while being grounded in established statistical methods.
+Despite its assumptions, such as linearity in the log-odds and independence of observations, logistic regression offers the flexibility to incorporate interaction terms and account for a range of predictor variables, making it a powerful tool for this research.
+
+##Model Validation
 
@@
 
+## Model Assumptions
+1. **Linearity**: The log odds of the outcome are linearly related to the predictors.
+2. **Independence**: Observations are independent of each other.
+3. **No Multicollinearity**: Predictors are not highly correlated.
+4. **No Outliers or Influential Points**: Checked through residual analysis.
+
 \newpage

### 2024-11-26 scripts/06-model_data.R
update model rds

@@ -1,37 +1,42 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Build a Logistic Regression Model to assess the likelihood of an individual not being married (outcome variable)
+# Author: Xizi Sun
+# Date: 24 November 2024
+# Contact: xizi.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: The `tidyverse`, `rstanarm`, `arrow` and `here` package must be installed
 
 
 #### Workspace setup ####
+# Load required libraries
 library(tidyverse)
 library(rstanarm)
+library(arrow)
+library(here)
 
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+#### Read the data and create model ####
+# Read the cleaned analysis dataset
+data <- read_parquet(here("data", "01-analysis_data", "analysis_data.parquet"))
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# Ensure the marital_status variable is converted to a binary response variable
+analysis_data <- data %>%
+  mutate(Not_Married = ifelse(marital_status == "Not_Married", 1, 0))
 
+# Fit a Bayesian logistic regression model
+bayesian_model <- stan_glm(
+  Not_Married ~ age + gender + Race + Income + education_level,
+  data = analysis_data,
+  family = binomial(link = "logit"),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  cores = 4,
+  adapt_delta = 0.99,
+  seed = 724
+)
 
 #### Save model ####
 saveRDS(
-  first_model,
[…]

### 2024-11-26 paper/paper.qmd
add model

@@
 library(knitr)
+library(broom)
+library(rstanarm)
+library(kableExtra)
+library(broom.mixed)
 ```
@@
 
-##Model Validation
+## Model Validation
+From @tbl-model-coefficients, the validation indicates that the logistic regression model provides a robust understanding of the predictors of "Not Married" status.
+However, some predictors, such as race, show variability in their impact, necessitating cautious interpretation.
+The model's good fit, coupled with well-defined confidence intervals and validation metrics, ensures that it provides reliable insights into the determinants of marital status.
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 
-# Results
+# Load the saved final model
+bayesian_model <-
+  readRDS(file = here::here("models/bayesian_model.rds"))
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-model-coefficients
+#| tbl-cap: Summary of the model
+
+# Extract coefficients
+coefficients <- broom::tidy(bayesian_model, conf.int = TRUE)
+coefficients |> kable(digits = 2)
+```
+
+## Results
 
 Our results are summarized in @tbl-modelresults.
+
+``````{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-coefficient
+#| fig-cap: Income Distribution for Unmarried Individuals
+
+# Preprocess the data to create a binary 'Not_Married' variable
+data$Not_Married <- ifelse(data$marital_status == "Not_Married", 1, 0)
+
+# Fit the logistic regression model
+model <- glm(Not_Married ~ age + gender + Race + Income + education_level, 
+             data = data, family = binomial())
+
+# Extract coefficients and confidence intervals
+model_summary <- broom::tidy(model, conf.int = TRUE)
+
+# Create the coefficient plot
+ggplot(model_summary, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+  geom_pointrange(color = "orange") +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  coord_flip() +
+  labs(
+    title = "Coefficients and Confidence Intervals from Logistic Regression",
+    x = "Predictors",
+    y = "Coefficient Value"
+  ) +
+  theme_minimal()
+```
 
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
 # Discussion

### 2024-11-26 paper/paper.qmd
update plots

@@
 
-Our results are summarized in @tbl-modelresults.
-
-``````{r}
+```{r}
 #| echo: false
@@
   labs(
-    title = "Coefficients and Confidence Intervals from Logistic Regression",
+    title = "Coefficients and Confidence Intervals",
     x = "Predictors",
@@
 ```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-edu-marriage
+#| fig-cap: Proportions of Marital Status by Education Level
+#| 
+# Calculate proportions of marital status within each education level
+education_marital_status <- data %>%
+  group_by(education_level, marital_status) %>%
+  summarise(count = n()) %>%
+  ungroup() %>%
+  group_by(education_level) %>%
+  mutate(proportion = count / sum(count))
 
+# Create the stacked bar chart with percentages and orange/blue colors
+ggplot(education_marital_status, aes(x = education_level, y = proportion, fill = marital_status)) +
+  geom_bar(stat = "identity", position = "stack") +
+  geom_text(
+    aes(label = scales::percent(proportion, accuracy = 1)),
+    position = position_stack(vjust = 0.5), size = 3.5, color = "white"
+  ) +
+  scale_fill_manual(values = c("Not_Married" = "#FFA500", "Married" = "#4682B4")) +
+  labs(
+    title = "Proportions of Marital Status by Education Level",
+    x = "Education Level",
+    y = "Proportion",
+    fill = "Marital Status"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-race-marriage
+#| fig-cap: Proportions of Marital Status by Race
+
+# Calculate proportions of marital status within each race
+race_marital_status <- data %>%
+  group_by(Race, marital_status) %>%
+  summarise(count = n()) %>%
+  ungroup() %>%
+  group_by(Race) %>%
+  mutate(proportion = count / sum(count))
+
+# Create the stacked bar chart with annotations
+ggplot(race_marital_status, aes(x = Race, y = proportion, fill = marital_status)) +
+  geom_bar(stat = "identity", position = "stack") +
+  geom_text(
+    aes(label = scales::percent(proportion, accuracy = 1)),
+    position = position_stack(vjust = 0.5), size = 3.5, color = "white"
+  ) +
+  scale_fill_manual(values = c("Not_Married" = "red", "Married" = "blue")) +
+  labs(
+    title = "Proportions of Marital Status by Race",
+    x = "Race",
+    y = "Proportion",
+    fill = "Marital Status"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-gender-marriage
+#| fig-cap: Marital Status Proportions by Gender
+
+# Calculate proportions of marital status within each gender
+gender_marital_status <- data %>%
+  group_by(gender, marital_status) %>%
+  summarise(count = n()) %>%
+  ungroup() %>%
+  group_by(gender) %>%
+  mutate(
+    proportion = count / sum(count),
+    percentage_label = paste0(scales::percent(proportion, accuracy = 1))
+  )
+
+# Separate pie charts for each gender with percentage labels
+ggplot(gender_marital_status, aes(x = "", y = proportion, fill = marital_status)) +
+  geom_bar(stat = "identity", width = 1) +
+  geom_text(
+    aes(label = percentage_label),
+    position = position_stack(vjust = 0.5), size = 4, color = "white"
+  ) +
+  coord_polar("y", start = 0) +
+  facet_wrap(~gender) +
+  scale_fill_manual(values = c("Not_Married" = "#FF9999", "Married" = "#66B3FF")) +
+  labs(
+    title = "Marital Status Proportions by Gender",
+    x = NULL,
+    y = NULL,
+    fill = "Marital Status"
+  ) +
+  theme_void() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16),
+    strip.text = element_text(size = 14),
+    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 10)
+  )
+```
 # Discussion

### 2024-11-26 paper/paper.qmd
update results

--- 2024-11-26 paper/paper.qmd
@@
 
-Our results are summarized in @tbl-modelresults.
+The analysis integrates multiple visualizations to provide a comprehensive understanding of the demographic, economic, and social dynamics associated with marital status.
+Below, we summarize the key findings derived from each visualization.
 
-``````{r}
+The logistic regression coefficients, as shown in @fig-coefficient, highlight the significant predictors of marital status.
+Race emerges as a crucial factor, with individuals identifying as "RaceBlack" or "RaceAsian" having substantial negative coefficients, indicating a reduced likelihood of being unmarried compared to other racial groups.
+Educational attainment also plays a major role; lower educational levels such as "Below High School" and "High School" are associated with a higher probability of not being married, as evidenced by their positive coefficients.
+Gender and income also contribute, with males marginally more likely to remain unmarried, while income has a smaller yet notable influence.
+
+```{r}
 #| echo: false
@@
 #| label: fig-coefficient
-#| fig-cap: Income Distribution for Unmarried Individuals
+#| fig-cap: Coefficients and Confidence Intervals
 
@@
   labs(
-    title = "Coefficients and Confidence Intervals from Logistic Regression",
     x = "Predictors",
@@
 
+@fig-Income-Unmarried depicts the income distribution of unmarried individuals, revealing that the majority of unmarried individuals fall into lower income brackets.
+The distribution is heavily skewed, with a significant concentration of individuals earning below $50,000 annually.
+The density sharply declines for higher income levels, reflecting the economic constraints that may influence marital status.
+A small proportion of high-income individuals remain unmarried, as shown by the long right tail of the distribution.
 ```{r}
@@
   geom_density(color = "blue", size = 1) +
-  ggtitle("Income Distribution for Unmarried Individuals") +
   xlab("Income") +
@@
 
+@fig-Age-Unmarried shows the age distribution for unmarried individuals, which is dominated by younger age groups, particularly those in their 20s and early 30s.
+The density decreases substantially as age increases, illustrating that as individuals age, the likelihood of remaining unmarried diminishes.
+This trend aligns with societal norms where marriage becomes more prevalent in middle age, reflecting the intersection of age-related societal expectations and personal decision-making.
 ```{r}
@@
   geom_density(color = "red", size = 1) +
-  ggtitle("Age Distribution for Unmarried Individuals") +
   xlab("Age") +
@@
 
+The proportions of marital status by education level are presented in @fig-edu-marriage.
+Individuals with higher education levels, such as "Above Bachelor" and "Bachelor," show significantly lower proportions of unmarried individuals, at 16% and 25%, respectively.
+Conversely, those with lower educational attainment, such as "Below High School" and "Some College," exhibit higher proportions of unmarried individuals, with 32% in each group.
+This suggests that educational attainment serves as a strong predictor of marital outcomes, potentially due to its influence on economic stability and social networks.
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-edu-marriage
+#| fig-cap: Proportions of Marital Status by Education Level
+#| 
+# Calculate proportions of marital status within each education level
+education_marital_status <- data %>%
+  group_by(education_level, marital_status) %>%
+  summarise(count = n()) %>%
+  ungroup() %>%
+  group_by(education_level) %>%
+  mutate(proportion = count / sum(count))
+
+# Create the stacked bar chart with percentages and orange/blue colors
+ggplot(education_marital_status, aes(x = education_level, y = proportion, fill = marital_status)) +
+  geom_bar(stat = "identity", position = "stack") +
+  geom_text(
+    aes(label = scales::percent(proportion, accuracy = 1)),
+    position = position_stack(vjust = 0.5), size = 3.5, color = "white"
+  ) +
+  scale_fill_manual(values = c("Not_Married" = "#FFA500", "Married" = "#4682B4")) +
+  labs(
+    x = "Education Level",
+    y = "Proportion",
+    fill = "Marital Status"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+@fig-race-marriage and @fig-gender-marriage highlight the proportions of marital status by race and gender.
+Race disparities are evident, with "American Indian" and "Black" populations showing higher proportions of unmarried individuals, at 37% and 43%, respectively.
+In contrast, "White" and "Asian" groups exhibit the lowest proportions of unmarried individuals, at 24% and 30%, respectively.
+Gender analysis reveals that males are more likely to remain unmarried than females, with 30% of males unmarried compared to 26% of females.
+These findings underscore the intersection of cultural, economic, and societal factors in shaping marital trends across both race and gender.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-race-marriage
+#| fig-cap: Proportions of Marital Status by Race
+
+# Calculate proportions of marital status within each race
+race_marital_status <- data %>%
+  group_by(Race, marital_status) %>%
+  summarise(count = n()) %>%
+  ungroup() %>%
+  group_by(Race) %>%
+  mutate(proportion = count / sum(count))
+
+# Create the stacked bar chart with annotations
+ggplot(race_marital_status, aes(x = Race, y = proportion, fill = marital_status)) +
+  geom_bar(stat = "identity", position = "stack") +
+  geom_text(
+    aes(label = scales::percent(proportion, accuracy = 1)),
+    position = position_stack(vjust = 0.5), size = 3.5, color = "white"
+  ) +
+  scale_fill_manual(values = c("Not_Married" = "red", "Married" = "blue")) +
+  labs(
+    x = "Race",
+    y = "Proportion",
+    fill = "Marital Status"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-gender-marriage
+#| fig-cap: Marital Status Proportions by Gender
+
+# Calculate proportions of marital status within each gender
+gender_marital_status <- data %>%
+  group_by(gender, marital_status) %>%
+  summarise(count = n()) %>%
+  ungroup() %>%
+  group_by(gender) %>%
+  mutate(
+    proportion = count / sum(count),
+    percentage_label = paste0(scales::percent(proportion, accuracy = 1))
+  )
+
+# Separate pie charts for each gender with percentage labels
+ggplot(gender_marital_status, aes(x = "", y = proportion, fill = marital_status)) +
+  geom_bar(stat = "identity", width = 1) +
+  geom_text(
+    aes(label = percentage_label),
+    position = position_stack(vjust = 0.5), size = 4, color = "white"
+  ) +
+  coord_polar("y", start = 0) +
+  facet_wrap(~gender) +
+  scale_fill_manual(values = c("Not_Married" = "#FF9999", "Married" = "#66B3FF")) +
+  labs(
+    x = NULL,
+    y = NULL,
+    fill = "Marital Status"
+  ) +
+  theme_void() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16),
+    strip.text = element_text(size = 14),
+    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 10)
+  )
+```
 # Discussion

### 2024-11-26 paper/paper.qmd
update paper

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Who Stays Single? Exploring the Factors Behind Non-Marriage in the United States"
+subtitle: "Low Income and Education Levels Correlate with Higher Propensity to Remain Unmarried"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study examines the socio-demographic determinants of non-marriage in the United States using logistic regression analysis on individual-level data.
+The findings highlight that individuals with lower income and education levels, particularly below a bachelor’s degree, are significantly more likely to remain unmarried.
+Additionally, gender and race also play critical roles, with men and certain racial groups, such as African Americans, showing higher unmarried rates.
+These results underscore structural inequities and socio-economic barriers shaping marital outcomes, offering insights into broader societal patterns and potential areas for intervention."
 format: pdf
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 
-Overview paragraph
+Marriage is often regarded as a cornerstone of social and economic stability, shaping the lives of individuals and the broader dynamics of society.
+However, in recent decades, the United States has experienced a notable shift in marital patterns, with an increasing proportion of the population remaining unmarried.
+This trend raises essential questions about the underlying factors contributing to the decision not to marry and the societal implications of such changes.
+Addressing these questions is vital for understanding broader socio-economic inequalities and cultural dynamics.
 
-Estimand paragraph
+This paper examines the determinants of non-marriage by analyzing a rich dataset from IPUMS USA [@ipumsusa], which offers harmonized census and survey data spanning demographic, social, and economic variables.
+Prior studies have largely focused on individual aspects of marital behavior, such as income or education, but few have comprehensively analyzed the intersection of multiple socio-demographic factors.
+This study bridges that gap by exploring the combined effects of education, income, race, gender, and age on marital outcomes.
 
-Results paragraph
+Using logistic regression analysis, the results reveal compelling patterns: individuals with higher education levels are more likely to marry, and economic stability, as indicated by higher income, is positively associated with marital status.
+Moreover, significant disparities are observed across racial and gender groups.
+African Americans exhibit higher non-marriage rates, while males show slightly elevated probabilities of remaining unmarried compared to females.
+These findings highlight persistent structural inequalities and cultural nuances shaping marriage in contemporary America.
 
-Why it matters paragraph
+Understanding the factors influencing non-marriage is crucial for addressing broader societal issues, including economic inequality, cultural shifts, and demographic changes.
+Marriage is often linked to financial stability, access to social networks, and overall well-being.
+The disparities uncovered in this study highlight systemic inequities that disproportionately affect certain groups, such as racial minorities and individuals with lower educational attainment.
+By identifying these disparities, policymakers and social institutions can better target interventions to reduce inequalities, promote inclusive opportunities for stable relationships, and adapt to the changing dynamics of family structures in modern society.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The remainder of this paper is organized as follows.
+@sec-data describes the dataset, variable selection, and measurement strategies used to examine marital status.
+@sec-model outlines the logistic regression approach, including the rationale behind model choice, prior specifications, and validation methods.
+@sec-results section presents key findings, supported by visualizations, and highlights the roles of education, income, race, gender, and age in predicting non-marriage.
+Finally, @sec-discussion explores the implications of these findings, acknowledges limitations, and proposes directions for future research.
 
@@
   labs(
-    title = "Proportion of Marital Status Categories",
     x = "Marital Status",
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-## Results
+## Results {#sec-results}
 
@@
 ```
-# Discussion
+# Discussion {#sec-discussion}
 
-## First discussion point {#sec-first-point}
+## Determinants of Non-Marriage and Societal Implications
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+This study delves into the socio-demographic factors that influence marital status, with a specific focus on the probability of remaining unmarried in the United States.
+By analyzing individual-level microdata through logistic regression, we identified significant predictors such as education, income, race, gender, and age.
 
-## Second discussion point
+For instance, the logistic regression model revealed that individuals with higher education levels, particularly those with a bachelor's degree or more, were significantly less likely to remain unmarried compared to those with lower education levels.
+Education, therefore, serves as a strong stabilizer in marital outcomes, providing individuals with economic resources and social capital that facilitate marriage.
+Simultaneously, the analysis underscored the heightened likelihood of remaining unmarried among African Americans, reflecting deeper structural inequities that impact marital patterns.
+Gender also emerged as a significant factor, with males exhibiting a slightly higher probability of being unmarried compared to females.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+These findings, grounded in data and supported by statistical rigor, affirm prior literature while adding layers of nuance.
+They emphasize how individual socio-demographic traits intersect with broader societal structures to shape marital outcomes.
 
-## Third discussion point
+## Insights into Education, Income, and Marriage Patterns
 
-## Weaknesses and next steps
+One of the most compelling findings of this study is the pronounced impact of education on marital status.
+Individuals with higher education levels—those with bachelor's degrees or above—demonstrated significantly lower rates of non-marriage.
+This is evident from the analysis showing that only 16% of individuals with education above a bachelor's degree were unmarried, compared to 32% of those with "Some College" education.
+The logistic regression coefficients also reinforced this relationship, with those holding higher degrees being far less likely to remain unmarried (e.g., coefficient for "Below High School" = 0.48).
 
-Weaknesses and next steps should also be included.
+This finding aligns with existing theories that higher education fosters economic stability, enhances social capital, and provides a greater sense of agency in life decisions, all of which are conducive to marriage.
+However, the results also highlighted that income, while less strongly predictive, plays a complementary role.
+The income distribution plot for unmarried individuals revealed clustering at lower income brackets, with the majority earning below $40,000 annually.
+This suggests that financial security continues to be an essential consideration in marital decisions, albeit to a lesser extent than education.
+
+These patterns point to the interplay between socio-economic resources and marital outcomes, emphasizing how disparities in access to education and income perpetuate inequalities in family formation.
+
+## Gender and Race in Marital Trends
+
+The analysis revealed striking disparities in marital trends across gender and racial lines.
+Gender-wise, men exhibited a higher likelihood of remaining unmarried compared to women, with 30% of men being unmarried versus 26% of women.
+This aligns with cultural norms and gendered expectations, where women are often perceived as more closely tied to family roles and marriage.
+Moreover, women may face greater societal pressure to marry, which could explain their lower rates of remaining unmarried.
+
+Racial disparities in marital status were even more pronounced.
+For example, African Americans exhibited the highest rates of non-marriage (43%), significantly higher than those for Asian Americans (30%) and White individuals (24%).
+These patterns likely reflect a combination of structural inequalities, cultural factors, and historical contexts.
+Economic disparities, higher unemployment rates, and systemic barriers to wealth accumulation among African Americans may contribute to these higher non-marriage rates.
+Conversely, Asian Americans, who exhibited the lowest rates of remaining unmarried, may benefit from cultural norms that strongly prioritize marriage and family cohesion.
+
+These findings underscore how structural inequities and cultural norms interact to shape marital trends.
+They also highlight the need for policies aimed at addressing systemic barriers and promoting equality in socio-economic outcomes to reduce disparities in family formation.
+
+## Weaknesses and Next Steps
+
+While this study provides critical insights into the predictors of marital status, it is not without limitations.
+The reliance on cross-sectional data, for example, restricts the ability to establish causal relationships.
+While the findings suggest strong associations between education, income, race, gender, and marital outcomes, longitudinal data could better capture how these factors interact over time to influence marriage decisions.
+
+Another limitation lies in the dataset itself, which does not include qualitative factors such as cultural attitudes, personal preferences, or psychological traits.
+These dimensions are critical for understanding the broader context of marital decisions.
+For instance, while income and education are robust predictors, societal expectations and individual life goals may play equally important roles that remain unquantified in this analysis.
+
+Future research should aim to integrate mixed-method approaches, combining quantitative rigor with qualitative depth, to provide a holistic view of marital trends.
+Additionally, while this study focused exclusively on the United States, extending the analysis to include international data could uncover global patterns and cultural nuances.
+For example, exploring how marriage trends differ in countries with varying economic systems, religious practices, and gender norms would provide valuable comparative insights.
+
+## Conclusion
+
+This study contributes to the understanding of non-marriage determinants by highlighting the interplay between education, income, race, gender, and age.
+Education emerged as the most significant predictor, with higher levels strongly associated with lower non-marriage rates.
+Income, while secondary, still played a notable role, emphasizing the importance of economic stability in marital decisions.
+Gendered expectations and racial disparities further revealed how societal structures and cultural norms influence marriage patterns.
+
+The findings highlight the socio-economic inequities that shape family formation and underscore the importance of addressing structural barriers to promote marital stability.
+Future research should continue to explore these dynamics, leveraging longitudinal data and integrating qualitative perspectives to deepen our understanding of marriage and its evolving role in contemporary society.
 

### 2024-11-26 paper/paper.qmd
update reference

@@
 Prior studies have largely focused on individual aspects of marital behavior, such as income or education, but few have comprehensively analyzed the intersection of multiple socio-demographic factors.
-This study bridges that gap by exploring the combined effects of education, income, race, gender, and age on marital outcomes.
+This study bridges that gap by exploring the combined effects of socio-demographic factors.
+Specifically, the estimand of this study is the reason of an individual remaining unmarried, conditioned on their socio-demographic attributes, including education level, income, race, gender, and age.
+This focus allows for a nuanced understanding of how these factors interact to influence marital decisions.
 
@@
 
-The dataset was analyzed using R [@citeR] and utilized several R packages for data manipulation and visualization, including ggplot2 [@citeGGplot2], dplyr [@citeDplyr], and here [@citeHere].
-The data was processed using the Apache Arrow package [@citeArrow] for efficient file handling and visualized with scales [@citeScales] for enhanced graphics.
-Reproducibility was ensured using knitr [@citeKnitr].
-Data was extracted from IPUMS USA [@ipumsusa].
-Guidance on storytelling with data was drawn from Telling Stories with Data [@tellingstories].
+The dataset was analyzed using R [@citeR] and utilized several R packages for data manipulation, visualization, and analysis, including ggplot2 [@ggplot2] for creating elegant graphics, dplyr [@citeDplyr] for data wrangling, and here [@citeHere] for simplifying file management.
+The data was processed efficiently using Apache Arrow [@citeArrow], and visualizations were further enhanced with the scales package [@citeScales].
+Bayesian modeling was implemented using rstanarm [@citeRstanarm], while statistical results were tidied with broom [@citeBroom] and broom.mixed [@citeBroomMixed].
+Tables were constructed using kableExtra [@citeKableExtra] for a polished presentation.
+Reproducibility was ensured with knitr [@citeKnitr].
+Data was extracted from IPUMS USA [@ipumsusa], and guidance on storytelling with data was drawn from Telling Stories with Data [@tellingstories].
 

### 2024-11-28 paper/paper.qmd
update charts in the paper

@@
 
-# Plot
-ggplot(marital_status_df, aes(x = Marital_Status, y = Proportion, fill = Marital_Status)) +
-  geom_bar(stat = "identity") +
+# Add percentage labels
+marital_status_df$Percentage <- paste0(round(marital_status_df$Proportion * 100, 1), "%")
+
+# Plot the pie chart with percentage labels
+ggplot(marital_status_df, aes(x = "", y = Proportion, fill = Marital_Status)) +
+  geom_bar(stat = "identity", width = 1) +
+  coord_polar(theta = "y") +
+  geom_text(aes(label = Percentage), position = position_stack(vjust = 0.5)) +
   labs(
-    x = "Marital Status",
-    y = "Proportion"
+    fill = "Marital Status",
+    title = "Distribution of Marital Status"
   ) +
-  theme_minimal()
+  theme_void()
 ```
@@
 #| message: false
-#| fig-cap: fig-predictor-variables
+#| fig-cap: Counts for Demographic Variables
 #| fig-subcap: ["Age Distribution", "Gender Distribution", "Race Distribution", "Education Level Distribution", "Income Distribution"]
@@
 #| label: fig-coefficient
-#| fig-cap: Coefficients and Confidence Intervals
+#| fig-cap: Coefficients of The Model
 

### 2024-12-03 scripts/06-model_data.R
update model files

@@ -23,7 +23,7 @@ analysis_data <- data %>%
   mutate(Not_Married = ifelse(marital_status == "Not_Married", 1, 0))
 
 # Fit a Bayesian logistic regression model
-bayesian_model <- stan_glm(
+logistic_model <- stan_glm(
   Not_Married ~ age + gender + Race + Income + education_level,
   data = analysis_data,
   family = binomial(link = "logit"),
@@ -36,7 +36,7 @@ bayesian_model <- stan_glm(
 
 #### Save model ####
 saveRDS(
-  bayesian_model,
-  file = "models/bayesian_model.rds"
+  logistic_model,
+  file = "models/logistic_model.rds"
 )
 

### 2024-12-03 other/datasheet/datasheet.qmd
update datasheet

---
title: "Datasheet for IPUMS USA Dataset"
author: 
  - Xizi Sun
thanks: "Code and data are available at: https://github.com/LilianS77/US_Marriage"
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from [@gebru2021datasheets].

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - This dataset was created to analyze the determinants of non-marriage in the United States. It addresses gaps in understanding how socio-demographic factors such as education, income, race, gender, and age influence marital status. Existing datasets lacked the level of detail and harmonization required for comprehensive modeling and analysis.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was curated using the Integrated Public Use Microdata Series (IPUMS USA), a project of the Minnesota Population Center.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
   - The IPUMS USA project is supported by the National Institutes of Health (NIH) and the National Science Foundation (NSF).
4. *Any other comments?*
    - No.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each instance in the dataset represents an individual in the United States, including demographic, economic, and social variables such as marital status, education level, race, and income.
2. *How many instances are there in total (of each type, if appropriate)?*
	- The dataset is based on a 1% national sample from the 2023 American Community Survey (ACS)
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- IPUMS USA's sampling strategy involves randomly generating a sample point for every hundred people in the census. The sampling strategy also ensures that dwellings have an equal probability of being included in the sample, regardless of their size.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each record includes raw data (e.g., income in dollars, age in years) and processed features (e.g., binary marital status: "Married" or "Not Married").
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- The target variable is marital status, categorized as "Married" or "Not Married."
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Missing data is minimal, with gaps primarily in income due to non-responses. The missing data are cleaned.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- The instances in the dataset are primarily independent individual records.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No specific splits are recommended for this dataset. However, for modeling purposes, researchers may opt to split the data into training, validation, and testing sets. A common practice would be an 80/10/10 split to ensure enough data for training while retaining sufficient records for validation and testing.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- The dataset has been harmonized by IPUMS USA, minimizing errors and redundancies. However, self-reported variables like income and marital status may introduce noise due to inaccuracies or social desirability bias. Additionally, missing data in income and education fields could affect analyses, though imputation methods have been used to mitigate this.

10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is self-contained and does not rely on external resources. It is available through the IPUMS USA platform, which provides archival access and ensures consistency over time. Usage is governed by the IPUMS user agreement, which prohibits redistribution and requires proper citation.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, the dataset does not contain confidential information. It is anonymized and complies with legal and ethical standards for data protection. Identifiable information, such as names or addresses, has been removed by IPUMS to ensure privacy.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- Yes, the dataset identifies sub-populations based on age, gender, race, income, and education level. These variables are explicitly labeled and allow for detailed demographic analyses.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No, the dataset is anonymized and does not allow direct or indirect identification of individuals. All personally identifiable information has been removed, and variables are aggregated or generalized to prevent re-identification.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- Yes, the dataset includes sensitive information such as race, income, and marital status. While these variables are anonymized and aggregated, they could still be sensitive due to their potential to reveal disparities or inequalities. No highly sensitive data, such as biometric information or government identifiers, is included.
16. *Any other comments?*
	- No.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was primarily collected through survey responses as part of the American Community Survey (ACS), conducted by the U.S. Census Bureau. These responses were directly provided by individuals in the sampled households. 
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The data was collected through a combination of mechanisms:
Mail: Respondents received surveys to complete and return.
Telephone: Follow-up interviews were conducted via phone for non-respondents.
Personal Visits: Trained enumerators conducted in-person interviews for households that did not respond via mail or phone.
Online Forms: Respondents could also complete the survey online.
Validation was ensured through rigorous protocols, including automated consistency checks, interviewer training, and manual quality reviews.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset is a 1% stratified random sample drawn from the larger ACS dataset. Stratification was based on demographic characteristics such as age, gender, race, and geographic location to ensure proportional representation of the U.S. population.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Data collection was carried out by the U.S. Census Bureau, involving trained enumerators, survey administrators, and data processing staff. Enumerators and survey administrators are federal employees compensated according to U.S. government pay scales. No external contractors or crowdworkers were involved in this process.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data for this study was collected as part of the 2023 American Community Survey. The ACS operates on a continuous basis, collecting data monthly throughout the year. The instances in this dataset reflect data collected from January to December 2023.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- The ACS and its data collection procedures are subject to review by the U.S. Census Bureau and adhere to federal laws and guidelines, including the Confidential Information Protection and Statistical Efficiency Act (CIPSEA). These processes ensure compliance with ethical standards for data collection, privacy protection, and data use.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was obtained from IPUMS USA, a third-party organization that harmonizes ACS data for research purposes. IPUMS USA receives data directly from the U.S. Census Bureau and applies its own harmonization procedures before making it available to researchers.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Yes, respondents were notified as part of the ACS process. The U.S. Census Bureau provides clear information about the purpose of the survey, the voluntary nature of participation, and how the data will be used. This information is communicated via mailed instructions, online forms, and during personal interviews.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Participation in the ACS is required by law under Title 13 of the United States Code. While respondents are informed about the mandatory nature of the survey, they are also assured that their data will be used strictly for statistical purposes and protected under federal confidentiality laws.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Since participation in the ACS is legally mandated, there is no mechanism for respondents to revoke their consent. However, individuals are not required to answer every question, and any unanswered questions are handled through imputation or estimation.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- The U.S. Census Bureau regularly conducts Privacy Impact Assessments (PIAs) to evaluate the risks associated with data collection and use. These assessments focus on maintaining respondent confidentiality and preventing misuse of data. IPUMS USA further ensures that datasets are anonymized and stripped of personally identifiable information before distribution.
12. *Any other comments?*
	- No.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Yes, preprocessing was performed on the raw dataset obtained from IPUMS USA to prepare it for analysis. The missing data are removed. The gender variable was cleaned to male and female. The education variable was grouped into five categories: "Below High School," "High School," "Some College," "Bachelor," and "Above Bachelor."
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- The raw data is downloaded from the IPUMS website and saved locally.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- The preprocessing was conducted using R, with packages such as dplyr for data manipulation and arrow for efficient file handling. The code for all preprocessing steps is included in the project's GitHub repository, providing full reproducibility.
4. *Any other comments?*
	- No.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	-  Yes, the dataset was used to analyze the socio-demographic determinants of non-marriage in the United States, focusing on factors such as education, income, race, gender, and age.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- Yes. Code and data are available at: https://github.com/LilianS77/US_Marriage.
3. *What (other) tasks could the dataset be used for?*
	- The dataset could be used for various tasks, including:
Studying income inequality and its impact on social behaviors.
Analyzing trends in educational attainment across different demographic groups.
Exploring intersections of race, gender, and economic outcomes
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- The dataset contains sensitive demographic variables such as race and income, which, if misused, could lead to stereotyping or biased interpretations. To mitigate these risks, users should follow ethical guidelines and contextualize findings appropriately.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset should not be used for:
Making deterministic conclusions about individuals.
Stereotyping or reinforcing biases against specific demographic groups.
Commercial purposes that violate the IPUMS terms of use.
6. *Any other comments?*
	- No

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- No, the dataset itself is not directly redistributed. Interested researchers can access the original data via IPUMS USA following their terms of use.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	-  The dataset is distributed through the IPUMS USA platform in formats such as CSV and Parquet. It does not have a specific DOI but is accessible via the IPUMS website.
3. *When will the dataset be distributed?*
	- The IPUMS USA data is updated annually, with the current dataset representing data collected in 2023
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Yes, the dataset is distributed under the IPUMS terms of use, which prohibit redistribution and require proper citation. There are no fees associated with academic use.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- Yes, the IPUMS terms of use restrict commercial use and redistribution of the dataset.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No export controls or regulatory restrictions apply. The dataset is anonymized and adheres to all U.S. legal standards for public use data.
7. *Any other comments?*
	- No.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The dataset is hosted and maintained by the Minnesota Population Center through the IPUMS USA platform.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- The dataset curators can be contacted via the IPUMS support email at ipums@umn.edu or through the support page on their website.
3. *Is there an erratum? If so, please provide a link or other access point.*
	-  IPUMS USA regularly updates its documentation and data to address potential issues. Any errata are published on the IPUMS USA website under the "Revisions and Errata" section.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Yes, the dataset is updated annually to include the latest ACS data. Updates are managed by IPUMS USA, and changes are communicated through the platform’s release notes and mailing list.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- The dataset contains anonymized data, and there are no specific retention limits since no personally identifiable information is included. The IPUMS platform ensures long-term availability for research purposes.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Older versions of the dataset are archived and accessible via the IPUMS platform. Researchers can access historical data spanning from 1850 to the most recent release to support longitudinal studies.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- While researchers cannot directly contribute to the IPUMS dataset, they can suggest improvements or report errors via the IPUMS feedback form. Extensions or augmentations must be managed independently and cannot be integrated into the IPUMS dataset due to its standardized format.
8. *Any other comments?*
	- No.

\newpage

# References

### 2024-12-03 scripts/06-model_data.R
update model

@@ -19,13 +19,13 @@ library(here)
 data <- read_parquet(here("data", "01-analysis_data", "analysis_data.parquet"))
 
 # Ensure the marital_status variable is converted to a binary response variable
-analysis_data <- data %>%
+data <- data %>%
   mutate(Not_Married = ifelse(marital_status == "Not_Married", 1, 0))
 
 # Fit a Bayesian logistic regression model
 logistic_model <- stan_glm(
   Not_Married ~ age + gender + Race + Income + education_level,
-  data = analysis_data,
+  data = data,
   family = binomial(link = "logit"),
   prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
   prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),

### 2024-12-03 paper/paper.qmd
update model section

@@
 ---
-title: "Who Stays Single? Exploring the Factors Behind Non-Marriage in the United States"
+title: "Who Stays Single? Exploring the Factors Behind Non-Marriage in the United States in 2023"
 subtitle: "Low Income and Education Levels Correlate with Higher Propensity to Remain Unmarried"
@@
 date-format: long
-abstract: "This study examines the socio-demographic determinants of non-marriage in the United States using logistic regression analysis on individual-level data.
+abstract: "This study examines the socio-demographic determinants of non-marriage in the United States in 2023 using logistic regression analysis on individual-level data.
 The findings highlight that individuals with lower income and education levels, particularly below a bachelor’s degree, are significantly more likely to remain unmarried.
@@
 library(broom.mixed)
+library(bayesplot)
 ```
@@
 
-Marriage is often regarded as a cornerstone of social and economic stability, shaping the lives of individuals and the broader dynamics of society.
-However, in recent decades, the United States has experienced a notable shift in marital patterns, with an increasing proportion of the population remaining unmarried.
-This trend raises essential questions about the underlying factors contributing to the decision not to marry and the societal implications of such changes.
-Addressing these questions is vital for understanding broader socio-economic inequalities and cultural dynamics.
+Marriage is widely perceived as a fundamental element of both social cohesion and economic stability.
+It profoundly influences individual lives and the broader social fabric.
+In recent decades, however, marital patterns in the United States have shifted noticeably, with a growing proportion of individuals choosing to remain unmarried.
+This trend prompts important questions about the factors underlying these decisions and the broader societal impacts of such changes.
+Understanding these issues is essential for addressing wider socio-economic inequalities and cultural transformations.
 
-This paper examines the determinants of non-marriage by analyzing a rich dataset from IPUMS USA [@ipumsusa], which offers harmonized census and survey data spanning demographic, social, and economic variables.
-Prior studies have largely focused on individual aspects of marital behavior, such as income or education, but few have comprehensively analyzed the intersection of multiple socio-demographic factors.
-This study bridges that gap by exploring the combined effects of socio-demographic factors.
-Specifically, the estimand of this study is the reason of an individual remaining unmarried, conditioned on their socio-demographic attributes, including education level, income, race, gender, and age.
-This focus allows for a nuanced understanding of how these factors interact to influence marital decisions.
+This study investigates the determinants of non-marriage by utilizing a dataset from IPUMS USA [@ipumsusa], which provides survey data encompassing demographic, social, and economic variables.
+While previous research has primarily examined individual factors like education or income in isolation, few studies have explored the interplay between multiple socio-demographic factors in shaping marital decisions.
+This research addresses this shortfall by analyzing these factors in combination.
 
-Using logistic regression analysis, the results reveal compelling patterns: individuals with higher education levels are more likely to marry, and economic stability, as indicated by higher income, is positively associated with marital status.
-Moreover, significant disparities are observed across racial and gender groups.
-African Americans exhibit higher non-marriage rates, while males show slightly elevated probabilities of remaining unmarried compared to females.
-These findings highlight persistent structural inequalities and cultural nuances shaping marriage in contemporary America.
+The estimand of this study is the reason an individual remains unmarried, analyzed in the context of their socio-demographic characteristics, including education level, income, race, gender, and age.
+This approach enables a deeper understanding of how these attributes collectively influence marital behavior.
 
-Understanding the factors influencing non-marriage is crucial for addressing broader societal issues, including economic inequality, cultural shifts, and demographic changes.
-Marriage is often linked to financial stability, access to social networks, and overall well-being.
-The disparities uncovered in this study highlight systemic inequities that disproportionately affect certain groups, such as racial minorities and individuals with lower educational attainment.
-By identifying these disparities, policymakers and social institutions can better target interventions to reduce inequalities, promote inclusive opportunities for stable relationships, and adapt to the changing dynamics of family structures in modern society.
+The findings, derived through logistic regression analysis, reveal significant patterns.
+Higher education levels are strongly linked to higher marriage rates, while economic stability, as indicated by higher income, positively correlates with marital status.
+Marked disparities are evident among racial and gender groups.
+African Americans, for instance, demonstrate higher rates of non-marriage, whereas males are slightly more likely to remain unmarried than females.
+These findings underscore persistent structural inequities and cultural complexities influencing marriage in contemporary American society.
+Recognizing these inequities equips policymakers and social institutions to design targeted interventions that address inequality, foster opportunities for stable relationships, and adapt to evolving family dynamics in modern society.
 
-The remainder of this paper is organized as follows.
-@sec-data describes the dataset, variable selection, and measurement strategies used to examine marital status.
-@sec-model outlines the logistic regression approach, including the rationale behind model choice, prior specifications, and validation methods.
-@sec-results section presents key findings, supported by visualizations, and highlights the roles of education, income, race, gender, and age in predicting non-marriage.
-Finally, @sec-discussion explores the implications of these findings, acknowledges limitations, and proposes directions for future research.
+The structure of this paper is as follows: @sec-data explains the dataset, variable selection, and measurement strategies employed to study marital status.
+@sec-model discusses the logistic regression methodology, including model rationale, prior specifications, and validation processes.
+@sec-results presents the key findings with supporting visualizations, emphasizing the roles of education, income, race, gender, and age in predicting non-marriage.
+Finally, @sec-discussion delves into the implications of the findings, acknowledges study limitations, and outlines potential areas for future research.
 
@@
 The data was processed efficiently using Apache Arrow [@citeArrow], and visualizations were further enhanced with the scales package [@citeScales].
-Bayesian modeling was implemented using rstanarm [@citeRstanarm], while statistical results were tidied with broom [@citeBroom] and broom.mixed [@citeBroomMixed].
+logistic regression modeling was implemented using rstanarm [@citeRstanarm], while statistical results were tidied with broom [@citeBroom] and broom.mixed [@citeBroomMixed].
 Tables were constructed using kableExtra [@citeKableExtra] for a polished presentation.
 Reproducibility was ensured with knitr [@citeKnitr].
+The bayesplot package [@citeBayesplot] was utilized for visualizing Bayesian model diagnostics and posterior predictive checks.
 Data was extracted from IPUMS USA [@ipumsusa], and guidance on storytelling with data was drawn from Telling Stories with Data [@tellingstories].
 
-## Data Source And Measurement
+## Measurement
 
-The data used for this study was sourced from **IPUMS USA (Integrated Public Use Microdata Series)** [@ipumsusa], a comprehensive repository providing access to harmonized census and survey data from the United States.
-IPUMS USA is renowned for its meticulous process of standardizing variables across datasets, ensuring compatibility over time, and providing extensive metadata to support research.
-This harmonization process enables researchers to conduct longitudinal and cross-sectional studies on social, demographic, and economic trends.
+For this study, data was specifically drawn from the American Community Survey (ACS) [@citeACS], a critical subset of IPUMS.
+The ACS conducts ongoing data collection throughout the year, selecting approximately 250,000 addresses each month to ensure the population's most current representation.
+The survey employs several collection methods to maximize participation.
+Initially, respondents receive a request to complete the survey online or return a paper questionnaire via mail.
+If no response is received, follow-up contact is made through telephone, utilizing a computer-assisted interview system (CATI).
+For those who remain unresponsive, a portion—approximately one-third—is selected for an in-person, computer-assisted personal interview (CAPI).
+In addition, the survey covers a wide range of social, economic and demographic characteristics, such as the variables of marital status, age and education analyzed in this study.
 
-The dataset consists of **individual-level microdata**, where each record represents a single person, numerically coded for all relevant characteristics.
-These records are organized into households, allowing for the study of individual behaviors and characteristics in the context of their family or co-residential settings.
-Unlike compiled statistics or pre-aggregated tables, this microdata structure provides researchers with unparalleled flexibility in exploring relationships between variables.
-
-To address the diversity of record layouts, coding schemes, and documentation across the historical scope of the dataset, IPUMS implements a rigorous **harmonization process**.
-Variables are assigned **uniform codes**, ensuring consistency across census years (1850–2010) and the American Community Surveys (ACS) (2000–present).
-This standardization simplifies the analysis of long-term trends and facilitates comparisons across time and space.
-
-## Variable Selection
-
-Using the IPUMS data extraction system, this study selected a focused subset of variables to examine the determinants of **non-marriage** in the United States.
-These include key demographic and socioeconomic characteristics, which are numerically coded for statistical analysis.
-
-The table [@tbl-cleaned-data] shows the variables after data cleaning,
-
-```{r}
-#| label: tbl-cleaned-data
-#| tbl-cap: Sample of cleaned data
-#| echo: false
-#| warning: false
-#| message: false
-
-# Display the first 6 rows of the dataset with column names
-head(data[, c("marital_status", "age", "gender", "Race", "Income", "education_level")], 6) |>
-  kable(
-    col.names = c("Marital Status", "Age", "Gender", "Race", "Income", "Education Level"),
-    booktabs = TRUE
-  )
-```
-
-### Outcome variables
+## Outcome Variables
 The primary outcome variable for this study is **Marital Status**, which categorizes individuals based on their marital state.
-The proportion of marital status categories is displayed in figure @fig-proportion-marriage.
+The proportion of marital status categories is displayed in @fig-proportion-marriage.
 This variable allows for a comparison between individuals who have never married (`Not_Married`) and those who have (`Married`).
@@
 
-### Predictor variables
+## Predictor Variables
 
-This section focuses on the predictor variables included in the study.
-The distribution of predictor variables is displayed in Figure @fig-predictor-variables.
+The distribution of predictor variables is displayed in @fig-predictor-variables-marital-status.
 These variables capture demographic, socioeconomic, and personal characteristics, providing a comprehensive framework for analyzing factors associated with marital status.
@@
 
+The unmarried results of the predictor variables are displayed in the @sec-results.
 ```{r}
@@
 #| message: false
-#| fig-cap: Counts for Demographic Variables
-#| fig-subcap: ["Age Distribution", "Gender Distribution", "Race Distribution", "Education Level Distribution", "Income Distribution"]
-#| label: fig-predictor-variables
+#| fig-cap: Counts for Demographic Variables by Marital Status
+#| fig-subcap: ["Age Distribution by Marital Status", "Gender Distribution by Marital Status", "Race Distribution by Marital Status", "Education Level Distribution by Marital Status", "Income Distribution by Marital Status"]
+#| label: fig-predictor-variables-marital-status
 #| layout-ncol: 2
 
-# Function to create a bar plot or histogram
-create_plot <- function(data, x_var, x_label, is_continuous = FALSE, binwidth = NULL) {
+# Function to create a bar plot or histogram with marital status differentiation
+create_plot_marital <- function(data, x_var, x_label, is_continuous = FALSE, binwidth = NULL) {
   if (is_continuous) {
-    # Create a histogram for continuous variables
-    p <- ggplot(data, aes_string(x = x_var)) +
-      geom_histogram(binwidth = binwidth, fill = "skyblue", color = "black") +
-      labs(x = x_label, y = "Count") +
+    # Create a histogram for continuous variables with marital status
+    p <- ggplot(data, aes_string(x = x_var, fill = "marital_status")) +
+      geom_histogram(binwidth = binwidth, position = "dodge", color = "black") +
+      labs(x = x_label, y = "Count", fill = "Marital Status") +
       theme_minimal() +
@@
   } else {
-    # Create a bar chart for categorical variables
-    p <- ggplot(data, aes_string(x = x_var)) +
-      geom_bar(fill = "skyblue", color = "black") +
-      labs(x = x_label, y = "Count") +
+    # Create a bar chart for categorical variables with marital status
+    p <- ggplot(data, aes_string(x = x_var, fill = "marital_status")) +
+      geom_bar(position = "dodge", color = "black") +
+      labs(x = x_label, y = "Count", fill = "Marital Status") +
       theme_minimal() +
@@
 
-# Create plots for the predictor variables
-create_plot(data, "age", "Age", is_continuous = TRUE, binwidth = 5)
-create_plot(data, "gender", "Gender")
-create_plot(data, "Race", "Race")
-create_plot(data, "education_level", "Education Level")
-create_plot(data, "Income", "Income", is_continuous = TRUE, binwidth = 10000)
-
+# Create plots for the predictor variables by marital status
+create_plot_marital(data, "age", "Age", is_continuous = TRUE, binwidth = 5)
+create_plot_marital(data, "gender", "Gender")
+create_plot_marital(data, "Race", "Race")
+create_plot_marital(data, "education_level", "Education Level")
+create_plot_marital(data, "Income", "Income", is_continuous = TRUE, binwidth = 10000)
 ```
@@
 
-To understand the phenomenon of non-marriage in the United States, I selected the IPUMS USA dataset over alternatives such as IPUMS International.
-While IPUMS International provides harmonized census data from 104 countries with over 1 billion person records, its vast scope makes it less targeted for this study.
-My focus is specifically on the U.S.
-population, and IPUMS USA offers high-precision data from American Community Surveys (ACS) and federal censuses, making it a more suitable choice for studying societal patterns specific to the United States.
+To examine the phenomenon of non-marriage in the United States, the IPUMS USA dataset was chosen as the primary data source, rather than IPUMS International.
+Although IPUMS International includes harmonized census data from 104 countries and encompasses over one billion individual records, its broad, global scope renders it less aligned with the objectives of this study.
+This research focuses specifically on societal patterns within the U.S.
+population.
+IPUMS USA, with its data derived from the American Community Surveys (ACS) and federal censuses, provides the precision and relevance needed to investigate non-marriage trends in the United States.
 
@@
 
-For this analysis, I employed a Logistic Regression Model to assess the likelihood of an individual not being married (outcome variable) based on several demographic and socioeconomic predictors (predictor variables).
+For this analysis, I apply a Logistic Regression Model to assess the likelihood of an individual not being married (outcome variable) based on several demographic and socioeconomic predictors (predictor variables).
 This model was chosen due to the binary nature of the outcome variable, which distinguishes between individuals who are "Not Married" versus those who are married (including divorced, widowed, or separated).
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
 
@@
 
-## Model justification
+## Model Justification
 
-The logistic regression model was chosen for this study due to its suitability for predicting binary outcomes, in this case, whether an individual is categorized as "Not Married" or not.
-Logistic regression is widely used in social science research to understand relationships between a binary dependent variable and multiple independent variables, making it an ideal choice for analyzing marital status.
-This model allows for a straightforward interpretation of the effects of predictors, such as age, income, education, and gender, through odds ratios, providing insight into the likelihood of being "Not Married." The dataset contains a mix of continuous and categorical variables, and logistic regression can handle this diversity effectively.
-Additionally, its theoretical simplicity ensures that the analysis remains interpretable while maintaining robustness.
-The choice of this model aligns with the study’s goal of identifying significant factors associated with non-marriage in the U.S., providing actionable insights while being grounded in established statistical methods.
-Despite its assumptions, such as linearity in the log-odds and independence of observations, logistic regression offers the flexibility to incorporate interaction terms and account for a range of predictor variables, making it a powerful tool for this research.
+The logistic regression model was chosen for this study due to its suitability in predicting binary outcomes.
+Specifically, it was employed to classify individuals as either "Not Married" or otherwise.
+This method is widely used in social science research to analyze the relationship between a binary dependent variable and several independent variables.
+It is therefore well-suited for examining patterns in marital status.
 
-## Model Validation
-From @tbl-model-coefficients, the validation indicates that the logistic regression model provides a robust understanding of the predictors of "Not Married" status.
-However, some predictors, such as race, show variability in their impact, necessitating cautious interpretation.
-The model's good fit, coupled with well-defined confidence intervals and validation metrics, ensures that it provides reliable insights into the determinants of marital status.
+The decisions regarding model design were guided by the dataset's structure and the variables under analysis.
+For instance, age was treated as a continuous variable to preserve its detail and avoid arbitrary groupings.
+Gender was modeled as a categorical variable with two levels—male and female—to reflect its binary nature in the data.
+Each racial category was included as a distinct level to capture the unique effects associated with different racial groups.
+Similarly, education and income were retained in their granular forms to maximize the use of the dataset’s available information.
+
+While the method depends on certain assumptions, such as linearity in the log-odds and independence of observations, it allows for the inclusion of interaction terms and accommodates a wide range of predictors.
+This balance of clarity, flexibility, and alignment with the study's objectives makes logistic regression an effective choice for this research.
+
+Posterior checks for the model and MCMC convergence check can be found in @sec-model-details.
+
+# Results {#sec-results}
+
+This study integrates multiple visualizations to provide a comprehensive understanding of the demographic, economic, and social dynamics associated with marital status.
+Below, we summarize the key findings derived from each visualization.
+
+@fig-coefficient and @tbl-model-summary highlight the coefficients for RaceWhite, RaceOther, and RaceAsian are negative, indicating that individuals from these racial groups are less likely to remain unmarried compared to others.
+The coefficient for RaceBlack is positive, and its confidence interval does not include 0, suggesting that RaceBlack group are more likely to remain unmarried compared to other racial groups.
+Individuals with lower levels of education (e.g., below high school) have a higher likelihood of remaining unmarried.
+The coefficient for males is positive, indicating that men are slightly more likely to remain unmarried compared to female.
 ```{r}
@@
 # Load the saved final model
-bayesian_model <-
-  readRDS(file = here::here("models/bayesian_model.rds"))
+logistic_model <-
+  readRDS(file = here::here("models/logistic_model.rds"))
 ```
@@
 #| message: false
-#| label: tbl-model-coefficients
+#| label: tbl-model-summary
 #| tbl-cap: Summary of the model
@@
 # Extract coefficients
-coefficients <- broom::tidy(bayesian_model, conf.int = TRUE)
+coefficients <- broom::tidy(logistic_model, conf.int = TRUE)
 coefficients |> kable(digits = 2)
 ```
-
-## Results {#sec-results}
-
-The analysis integrates multiple visualizations to provide a comprehensive understanding of the demographic, economic, and social dynamics associated with marital status.
-Below, we summarize the key findings derived from each visualization.
-
-The logistic regression coefficients, as shown in @fig-coefficient, highlight the significant predictors of marital status.
-Race emerges as a crucial factor, with individuals identifying as "RaceBlack" or "RaceAsian" having substantial negative coefficients, indicating a reduced likelihood of being unmarried compared to other racial groups.
-Educational attainment also plays a major role; lower educational levels such as "Below High School" and "High School" are associated with a higher probability of not being married, as evidenced by their positive coefficients.
-Gender and income also contribute, with males marginally more likely to remain unmarried, while income has a smaller yet notable influence.
 
@@
 The density decreases substantially as age increases, illustrating that as individuals age, the likelihood of remaining unmarried diminishes.
-This trend aligns with societal norms where marriage becomes more prevalent in middle age, reflecting the intersection of age-related societal expectations and personal decision-making.
 ```{r}
@@
 The proportions of marital status by education level are presented in @fig-edu-marriage.
-Individuals with higher education levels, such as "Above Bachelor" and "Bachelor," show significantly lower proportions of unmarried individuals, at 16% and 25%, respectively.
-Conversely, those with lower educational attainment, such as "Below High School" and "Some College," exhibit higher proportions of unmarried individuals, with 32% in each group.
-This suggests that educational attainment serves as a strong predictor of marital outcomes, potentially due to its influence on economic stability and social networks.
+Individuals with higher education levels show significantly lower proportions of unmarried individuals, with 25% for those holding a bachelor's degree and 16% for those with education above a bachelor's degree.
+In contrast, individuals with lower educational attainment exhibit higher proportions of unmarried individuals, with 32%, 31%, and 32% for those with below high school, high school, and some college education, respectively.
 ```{r}
@@
 Gender analysis reveals that males are more likely to remain unmarried than females, with 30% of males unmarried compared to 26% of females.
-These findings underscore the intersection of cultural, economic, and societal factors in shaping marital trends across both race and gender.
 
@@
 ```
+
 ```{r}
@@
 
-For instance, the logistic regression model revealed that individuals with higher education levels, particularly those with a bachelor's degree or more, were significantly less likely to remain unmarried compared to those with lower education levels.
+For instance, the logistic regression model revealed that individuals with higher education levels, particularly those with a bachelor's degree or above, were significantly less likely to remain unmarried compared to those with lower education levels.
 Education, therefore, serves as a strong stabilizer in marital outcomes, providing individuals with economic resources and social capital that facilitate marriage.
@@
 
-## How to Extract 2022 ACS Data from IPUMS
+## Data Details {#sec-data-details}
 
-To obtain data from IPUMS, we start by navigating to the IPUMS USA section and clicking on Get Data.
-Next, we go to the Select Sample section, where we uncheck the "Default sample from each year" option and instead select 2023 ACS.
-After selecting our sample, we proceed to add variables of interest.
-For individual-level data, we might add variables from the Person section.
-For example, under Demographic, we could include variables like AGE, and under Person, we could add SEX, RACE, INCTOT (total personal income) and EDUC (education attainment).
-Once our variables are selected, we click View Cart, then proceed by clicking Create Data Extract.
-At this point, we review our selections, change the Data Format to CSV, and submit our extract for processing.
-Then we saved it locally as usa_00001.csv.
+### Cleaned Data
 
-## Model Assumptions
-1. **Linearity**: The log odds of the outcome are linearly related to the predictors.
-2. **Independence**: Observations are independent of each other.
-3. **No Multicollinearity**: Predictors are not highly correlated.
-4. **No Outliers or Influential Points**: Checked through residual analysis.
+The table [@tbl-cleaned-data] shows the analysis data after data cleaning.
+
+```{r}
+#| label: tbl-cleaned-data
+#| tbl-cap: Sample of cleaned data
+#| echo: false
+#| warning: false
+#| message: false
+
+# Display the first 6 rows of the dataset with column names
+head(data[, c("marital_status", "age", "gender", "Race", "Income", "education_level")], 6) |>
+  kable(
+    col.names = c("Marital Status", "Age", "Gender", "Race", "Income", "Education Level"),
+    booktabs = TRUE
+  )
+```
+
+## Model Details {#sec-model-details}
+
+### Posterior predictive check
+
+@fig-post provides an illustration of the posterior predictive check, showing the alignment between the observed outcome variable and the simulations derived from the posterior distribution.
+In a similar vein, @fig-prior-post depicts a comparison between the prior and posterior distributions, highlighting the degree to which the data informs and updates our estimates.
+Together, these figures offer strong evidence that the model achieves a reliable and accurate fit to the observed data.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-post
+#| fig-cap: Posterior distribution for logistic regression model
+
+pp_check(logistic_model) +
+  theme(legend.position = "bottom")
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-prior-post
+#| fig-cap: Comparing prior distribution with posterior distribution
+
+# Compare posterior and prior distributions
+posterior_vs_prior(logistic_model) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(hjust = 1, vjust = 0.5),
+    text = element_text(size = 8)
+  ) +
+  guides(color = guide_legend(ncol = 6)) +
+  theme(legend.position = "bottom") +
+  coord_flip() +
+  ggtitle("Comparison of Prior and Posterior Distributions")
+```
+
+### Diagnostics
+
+@fig-trace1 and @fig-rhat provide insights into the MCMC diagnostics for the logistic regression model.
+@fig-trace1 show stable and well-mixed chains for all parameters, indicating convergence.
+fig-rhat shows $\hat{R}$ diagnostics are all close to 1, further confirming that the MCMC sampling has successfully converged to the posterior distribution.
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-trace1
+#| fig-cap: "MCMC Convergence Check: Trace Plots for Key Parameters"
+#| fig-subcap: ["Intercept", "genderMale", "education_levelBachelor", "education_levelHigh_School", "education_levelBelow_High_School", "education_levelSome_College", "RaceAsian", "RaceBlack", "RaceOther", "RaceWhite", "age", "Income"]
+
+# Trace plots for select parameters
+library(bayesplot)
+mcmc_trace(
+  as.array(logistic_model), # Replace `model_fit` with your fitted model object
+  pars = c("(Intercept)", "genderMale", "education_levelBachelor", 
+           "education_levelHigh_School", "education_levelBelow_High_School", 
+           "education_levelSome_College", "RaceAsian", "RaceBlack", 
+           "RaceOther", "RaceWhite", "age", "Income")
+)+
+  scale_x_continuous(
+    breaks = seq(0, 1000, by = 200), # Set x-axis breaks at intervals of 200
+    labels = seq(0, 1000, by = 200)  # Match labels to breaks
+  ) +
+  theme_minimal() +
+  theme(
+    text = element_text(size = 10), # Adjust text size for clarity
+    legend.position = "bottom"      # Move legend to the bottom
+  )
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-rhat
+#| fig-cap: "Rhat plot"
+
+plot(logistic_model, "rhat")
+```
+## Surveys, sampling, and observational data {#sec-survey-sampling-observation}
+
+### Idealized Methodology
+
+#### Overview
+
+The objective of this survey is to explore the socio-demographic determinants of non-marriage in the United States.
+The study aims to understand factors influencing decisions to remain unmarried, including education, income, race, age, and cultural attitudes.
+A budget of $50,000 is allocated to collect diverse and representative data from individuals aged 18 and above.
+The methodology is designed to maximize the accuracy.
+
+#### Sampling Approach
+
+A stratified sampling approach is adopted to ensure the representation of key demographic and socio-economic groups. 
+Stratification variables include:
+
+* Age: 18–29, 30–44, 45–64, 65+
+* Gender: Male, Female, Non-binary
+* Race/Ethnicity: White, Black, Hispanic, Asian, American Indian, Other
+* Income: <$20,000, $20,000–$59,999, $60,000–$99,999, >$100,000
+* Education: Under High School, High School, Some College, Bachelor's Degree, Above Bachelor's Degree
+*Geographic Region: Northeast, Midwest, South, West
+The target sample size is 5,000 respondents, yielding a margin of error of ±2.5% at a 95% confidence level.
+
+#### Recruitment Strategy
+
+* Online Surveys: Partner with established survey platforms like Prolific or Qualtrics, which offer access to diverse and validated panels. 
+* Community Engagement: Partnerships with community organizations for in-person recruitment.
+* Mail Invitations: Surveys sent by mail, accompanied by pre-paid return envelopes or online access codes for respondents to complete the survey.
+* Telephone Interviews: Random-digit dialing (RDD) to reach both landlines and mobile phones, ensuring coverage of participants without internet access.
+
+#### Data Collection
+
+Data collection will utilize platforms like Google Forms, with telephone and mail surveys structured to mirror the design and flow of the Google Form questionnaire.
+To reduce the likelihood of participants abandoning the survey, it has been tailored to take approximately five to ten minutes to complete.
+
+#### Data Validation and Quality Control
+
+* Pre-Survey Validation: Pilot testing with diverse focus groups to refine question phrasing and structure.
+* Ongoing Quality Checks: Automated checks for inconsistent or incomplete responses during the survey.
+* Post-Survey Weighting: Application of population weights to correct for sample imbalances across demographic strata.
+
+#### Budget Allocation
+
+* Survey Recruitment: $30,000
+* Incentives: $10,000
+* Data Cleaning and Analysis: $5,000
+* Miscellaneous (software, outreach): $5,000
+
+### Idealized Survey Questions
+
+Welcome Message  
+Welcome!
+Thank you for participating in this survey on marital status and the factors influencing people's decision to remain unmarried.
+Your responses are anonymous and will only be used for research purposes.
+This survey will take approximately 5–10 minutes to complete.
+
+---
+
+Screening Questions  
+Q1: Are you currently residing in the United States?  
+- Yes  
+- No *(If "No," terminate the survey with a thank-you message.)*
+
+Q2: What is your age?  
+- Under 18 *(Terminate the survey with a thank-you message.)*  
+- 18 or older
+
+---
+
+Consent to Participate  
+Before proceeding, please read the following:  
+- Your participation in this survey is voluntary.  
+- You may skip any question or withdraw at any time.  
+- All responses are anonymous and will be aggregated for research purposes.
+
+Do you agree to participate in this survey?  
+- Yes, I agree to participate.  
+- No, I do not agree to participate. *(If "No," terminate the survey with a thank-you message.)*
+
+---
+
+Survey Questions
+
+Q3: What is your gender?  
+- Male  
+- Female  
+- Non-binary  
+- Prefer not to say
+
+Q4: What is your race/ethnicity?  
+- White  
+- Black or African American  
+- Asian  
+- Hispanic or Latino  
+- Native American or Alaska Native  
+- Other
+
+Q5: What is your highest level of education?  
+- Less than high school  
+- High school diploma or equivalent  
+- Some college  
+- Bachelor’s degree  
+- Postgraduate degree
+
+Q6: What is your total annual income?  
+- Less than $20,000  
+- $20,000–$49,999  
+- $50,000–$79,999  
+- $80,000–$99,999  
+- $100,000 or more
+
+Q7: What is your age group?  
+- 18–29  
+- 30–44  
+- 45–59  
+- 60 or older
+
+---
+
+Q8: What is your current marital status?  
+- Never married  
+- Married  
+- Divorced  
+- Widowed  
+- Separated
+
+Q9: If never married, have you ever been in a serious relationship (e.g., cohabitation, long-term dating)?  
+- Yes  
+- No
+
+---
+
+Q10: How important do you consider marriage in your life?  
+- Very important  
+- Somewhat important  
+- Not very important  
+- Not at all important
+
+Q11: What are the main reasons for not being married? *(Select all that apply)*  
+- Financial concerns  
+- Lack of a suitable partner  
+- Personal choice  
+- Career priorities  
+- Family or cultural pressures  
+- Other (please specify)
+
+Q12: Do you feel societal pressure to get married?  
+- Strongly agree  
+- Agree  
+- Neutral  
+- Disagree  
+- Strongly disagree
+
+---
+
+Q13: Do you plan to get married in the future?  
+- Yes  
+- No  
+- Unsure
+
+Q14: If unsure or no, what factors might influence your decision?  
+- Improved financial stability  
+- Finding the right partner  
+- Change in personal values  
+- Other (please specify)
+
+---
+
+Closing Message  
+Thank you for completing this survey! Your responses have been recorded. If you have any questions about this survey or its purpose, please contact xizi.sun@mail.utoronto.ca. Have a great day!
 

### 2024-12-03 paper/paper.qmd
final commit

@@
 The findings highlight that individuals with lower income and education levels, particularly below a bachelor’s degree, are significantly more likely to remain unmarried.
-Additionally, gender and race also play critical roles, with men and certain racial groups, such as African Americans, showing higher unmarried rates.
+Additionally, gender and race also play important roles, with men and certain racial groups, such as African Americans, showing higher unmarried rates.
 These results underscore structural inequities and socio-economic barriers shaping marital outcomes, offering insights into broader societal patterns and potential areas for intervention."
@@
 
-For this study, data was specifically drawn from the American Community Survey (ACS) [@citeACS], a critical subset of IPUMS.
+For this study, data was specifically drawn from the American Community Survey (ACS) [@citeACS], a subset of IPUMS.
 The ACS conducts ongoing data collection throughout the year, selecting approximately 250,000 addresses each month to ensure the population's most current representation.
@@
     p <- ggplot(data, aes_string(x = x_var, fill = "marital_status")) +
-      geom_histogram(binwidth = binwidth, position = "dodge", color = "black") +
+      geom_histogram(binwidth = binwidth, position = "dodge") +
       labs(x = x_label, y = "Count", fill = "Marital Status") +
@@
     p <- ggplot(data, aes_string(x = x_var, fill = "marital_status")) +
-      geom_bar(position = "dodge", color = "black") +
+      geom_bar(position = "dodge") +
       labs(x = x_label, y = "Count", fill = "Marital Status") +
@@
 
-This study delves into the socio-demographic factors that influence marital status, with a specific focus on the probability of remaining unmarried in the United States.
-By analyzing individual-level microdata through logistic regression, we identified significant predictors such as education, income, race, gender, and age.
+This study investigates the socio-demographic factors influencing marital status, focusing on the probability of remaining unmarried in the United States.
+Through logistic regression analysis applied to individual-level microdata, significant predictors such as education, income, race, gender, and age were identified.
 
-For instance, the logistic regression model revealed that individuals with higher education levels, particularly those with a bachelor's degree or above, were significantly less likely to remain unmarried compared to those with lower education levels.
-Education, therefore, serves as a strong stabilizer in marital outcomes, providing individuals with economic resources and social capital that facilitate marriage.
-Simultaneously, the analysis underscored the heightened likelihood of remaining unmarried among African Americans, reflecting deeper structural inequities that impact marital patterns.
-Gender also emerged as a significant factor, with males exhibiting a slightly higher probability of being unmarried compared to females.
+Higher education levels were strongly associated with lower non-marriage rates.
+For instance, Figure @fig-edu-marriage shows that only 16% of individuals with education above a bachelor’s degree were unmarried, compared to 32% for those with "Some College" education.
+Logistic regression coefficients further underscore this trend, with individuals holding higher degrees being substantially less likely to remain unmarried (e.g., the coefficient for "Below High School" = 0.48).
+Education thus acts as a stabilizing factor, equipping individuals with economic resources and social networks conducive to marriage.
 
-These findings, grounded in data and supported by statistical rigor, affirm prior literature while adding layers of nuance.
-They emphasize how individual socio-demographic traits intersect with broader societal structures to shape marital outcomes.
+Similarly, income plays a complementary but less pronounced role.
+@fig-income-distribution illustrates that unmarried individuals tend to cluster in lower income brackets, with the majority earning less than $40,000 annually.
+This highlights the persistent importance of financial stability in marital decisions, although its impact appears secondary to education.
+
+The results also emphasize significant disparities across racial and gender lines.
+African Americans exhibited the highest non-marriage rates (43%), as shown in @fig-race-marriage, significantly surpassing those of Asian Americans (30%) and White individuals (24%).
+These patterns likely reflect structural inequalities, such as higher unemployment rates and barriers to wealth accumulation among African Americans.
+Gender differences were also evident, with men showing a slightly higher likelihood of remaining unmarried compared to women (30% versus 26%, respectively, @fig-gender-marriage).
+This aligns with cultural norms where women are more closely tied to family roles and experience greater societal pressure to marry.
+
+These findings reinforce how individual socio-demographic traits intersect with societal structures to shape marital outcomes.
+They also highlight systemic inequities that warrant targeted interventions.
 
@@
 
-One of the most compelling findings of this study is the pronounced impact of education on marital status.
-Individuals with higher education levels—those with bachelor's degrees or above—demonstrated significantly lower rates of non-marriage.
-This is evident from the analysis showing that only 16% of individuals with education above a bachelor's degree were unmarried, compared to 32% of those with "Some College" education.
-The logistic regression coefficients also reinforced this relationship, with those holding higher degrees being far less likely to remain unmarried (e.g., coefficient for "Below High School" = 0.48).
+Education emerged as the most significant predictor in this analysis.
+Higher education levels correlate strongly with lower non-marriage rates, as depicted in @fig-edu-marriage and @fig-coefficient.
+These findings align with theories suggesting that education enhances economic stability, expands social capital, and provides greater agency in life choices, all of which promote marriage.
+However, as @fig-income-distribution indicates, financial stability remains an essential but secondary factor in marital decisions.
 
-This finding aligns with existing theories that higher education fosters economic stability, enhances social capital, and provides a greater sense of agency in life decisions, all of which are conducive to marriage.
-However, the results also highlighted that income, while less strongly predictive, plays a complementary role.
-The income distribution plot for unmarried individuals revealed clustering at lower income brackets, with the majority earning below $40,000 annually.
-This suggests that financial security continues to be an essential consideration in marital decisions, albeit to a lesser extent than education.
-
-These patterns point to the interplay between socio-economic resources and marital outcomes, emphasizing how disparities in access to education and income perpetuate inequalities in family formation.
+The interplay between education and income underscores broader socio-economic disparities.
+For example, individuals with lower education levels not only face limited economic opportunities but are also more likely to remain unmarried, perpetuating cycles of inequality in family formation.
 
@@
 
-The analysis revealed striking disparities in marital trends across gender and racial lines.
-Gender-wise, men exhibited a higher likelihood of remaining unmarried compared to women, with 30% of men being unmarried versus 26% of women.
-This aligns with cultural norms and gendered expectations, where women are often perceived as more closely tied to family roles and marriage.
-Moreover, women may face greater societal pressure to marry, which could explain their lower rates of remaining unmarried.
+The analysis revealed gender and racial disparities.
+As shown in @fig-gender-marriage, men exhibit higher non-marriage rates compared to women, likely reflecting cultural expectations and differing societal pressures.
+Racial disparities, however, are even more striking.
+African Americans, for instance, show the highest rates of non-marriage, suggesting systemic inequities such as income disparities and historical barriers to wealth accumulation.
+By contrast, Asian Americans, with the lowest non-marriage rates, may benefit from cultural norms emphasizing family cohesion and marriage stability.
 
-Racial disparities in marital status were even more pronounced.
-For example, African Americans exhibited the highest rates of non-marriage (43%), significantly higher than those for Asian Americans (30%) and White individuals (24%).
-These patterns likely reflect a combination of structural inequalities, cultural factors, and historical contexts.
-Economic disparities, higher unemployment rates, and systemic barriers to wealth accumulation among African Americans may contribute to these higher non-marriage rates.
-Conversely, Asian Americans, who exhibited the lowest rates of remaining unmarried, may benefit from cultural norms that strongly prioritize marriage and family cohesion.
-
-These findings underscore how structural inequities and cultural norms interact to shape marital trends.
-They also highlight the need for policies aimed at addressing systemic barriers and promoting equality in socio-economic outcomes to reduce disparities in family formation.
+These patterns illustrate how cultural norms and structural inequities converge to shape marital outcomes.
+Addressing these disparities requires policies that promote economic equity and social inclusivity.
 
@@
 
-While this study provides critical insights into the predictors of marital status, it is not without limitations.
-The reliance on cross-sectional data, for example, restricts the ability to establish causal relationships.
-While the findings suggest strong associations between education, income, race, gender, and marital outcomes, longitudinal data could better capture how these factors interact over time to influence marriage decisions.
+While this study provides valuable insights into the determinants of non-marriage, it is not without limitations.
+The cross-sectional nature of the data restricts causal inferences, making it difficult to ascertain how variables such as education and income evolve over time to influence marital decisions.
+Longitudinal data would allow for a deeper exploration of these dynamics.
 
-Another limitation lies in the dataset itself, which does not include qualitative factors such as cultural attitudes, personal preferences, or psychological traits.
-These dimensions are critical for understanding the broader context of marital decisions.
-For instance, while income and education are robust predictors, societal expectations and individual life goals may play equally important roles that remain unquantified in this analysis.
+Additionally, the dataset lacks qualitative dimensions, such as cultural attitudes or personal preferences, which are crucial for understanding the broader context of marital decisions.
+For instance, while financial stability and education are strong predictors, societal expectations and individual life goals may play equally important roles that remain unmeasured.
 
-Future research should aim to integrate mixed-method approaches, combining quantitative rigor with qualitative depth, to provide a holistic view of marital trends.
-Additionally, while this study focused exclusively on the United States, extending the analysis to include international data could uncover global patterns and cultural nuances.
-For example, exploring how marriage trends differ in countries with varying economic systems, religious practices, and gender norms would provide valuable comparative insights.
+Future research should consider integrating mixed-method approaches to address these gaps.
+Furthermore, expanding the analysis to include international datasets could offer comparative insights, shedding light on how cultural and economic systems influence marriage patterns globally.
 
@@
 
-This study contributes to the understanding of non-marriage determinants by highlighting the interplay between education, income, race, gender, and age.
+This study contributes to the understanding of non-marriage by emphasizing the roles of education, income, race, gender, and age.
 Education emerged as the most significant predictor, with higher levels strongly associated with lower non-marriage rates.
-Income, while secondary, still played a notable role, emphasizing the importance of economic stability in marital decisions.
-Gendered expectations and racial disparities further revealed how societal structures and cultural norms influence marriage patterns.
+Income, while secondary, remains a critical factor in promoting marital stability.
+Racial and gender disparities further highlight systemic inequities that shape marital trends.
 
-The findings highlight the socio-economic inequities that shape family formation and underscore the importance of addressing structural barriers to promote marital stability.
-Future research should continue to explore these dynamics, leveraging longitudinal data and integrating qualitative perspectives to deepen our understanding of marriage and its evolving role in contemporary society.
+By addressing these structural barriers, policymakers and institutions can promote equity in family formation.
+Future research should build on these findings, leveraging longitudinal and qualitative approaches to provide a more comprehensive understanding of marriage and its evolving role in contemporary society.
 

### 2024-12-03 paper/paper.qmd
fix the sentences

@@
 Higher education levels were strongly associated with lower non-marriage rates.
-For instance, Figure @fig-edu-marriage shows that only 16% of individuals with education above a bachelor’s degree were unmarried, compared to 32% for those with "Some College" education.
+For instance, @fig-edu-marriage shows that only 16% of individuals with education above a bachelor’s degree were unmarried, compared to 32% for those with "Some College" education.
 Logistic regression coefficients further underscore this trend, with individuals holding higher degrees being substantially less likely to remain unmarried (e.g., the coefficient for "Below High School" = 0.48).
@@
 These findings align with theories suggesting that education enhances economic stability, expands social capital, and provides greater agency in life choices, all of which promote marriage.
-However, as @fig-income-distribution indicates, financial stability remains an essential but secondary factor in marital decisions.
+However, as @fig-Income-Unmarried indicates, financial stability remains an essential but secondary factor in marital decisions.
 
@@
 
-The table [@tbl-cleaned-data] shows the analysis data after data cleaning.
+The @tbl-cleaned-data shows the analysis data after data cleaning.
 

### 2024-12-03 paper/paper.qmd
Final commit

@@
 Similarly, income plays a complementary but less pronounced role.
-@fig-income-distribution illustrates that unmarried individuals tend to cluster in lower income brackets, with the majority earning less than $40,000 annually.
+@fig-Income-Unmarried illustrates that unmarried individuals tend to cluster in lower income brackets, with the majority earning less than $40,000 annually.
 This highlights the persistent importance of financial stability in marital decisions, although its impact appears secondary to education.

### 2024-12-03 paper/paper.qmd
update paper

@@
 Marriage is widely perceived as a fundamental element of both social cohesion and economic stability.
-It profoundly influences individual lives and the broader social fabric.
+It greatly influences individual lives and the broader social fabric.
 In recent decades, however, marital patterns in the United States have shifted noticeably, with a growing proportion of individuals choosing to remain unmarried.
@@
 
-The findings, derived through logistic regression analysis, reveal significant patterns.
+The findings, derived through logistic regression analysis, show significant patterns.
 Higher education levels are strongly linked to higher marriage rates, while economic stability, as indicated by higher income, positively correlates with marital status.
@@
 @sec-results presents the key findings with supporting visualizations, emphasizing the roles of education, income, race, gender, and age in predicting non-marriage.
-Finally, @sec-discussion delves into the implications of the findings, acknowledges study limitations, and outlines potential areas for future research.
+Finally, @sec-discussion discusses the implications of the findings, acknowledges study limitations, and outlines potential areas for future research.
 
@@
 The distribution of predictor variables is displayed in @fig-predictor-variables-marital-status.
-These variables capture demographic, socioeconomic, and personal characteristics, providing a comprehensive framework for analyzing factors associated with marital status.
+These variables capture demographic, socioeconomic, and personal characteristics, providing a thorough framework for analyzing factors associated with marital status.
 Below are the key predictor variables:
@@
 
-This study integrates multiple visualizations to provide a comprehensive understanding of the demographic, economic, and social dynamics associated with marital status.
+This study integrates multiple visualizations to provide a thorough understanding of the demographic, economic, and social dynamics associated with marital status.
 Below, we summarize the key findings derived from each visualization.
@@
 
-@fig-Income-Unmarried depicts the income distribution of unmarried individuals, revealing that the majority of unmarried individuals fall into lower income brackets.
+@fig-Income-Unmarried depicts the income distribution of unmarried individuals, showing that the majority of unmarried individuals fall into lower income brackets.
 The distribution is heavily skewed, with a significant concentration of individuals earning below $50,000 annually.
@@
 In contrast, "White" and "Asian" groups exhibit the lowest proportions of unmarried individuals, at 24% and 30%, respectively.
-Gender analysis reveals that males are more likely to remain unmarried than females, with 30% of males unmarried compared to 26% of females.
+Gender analysis shows that males are more likely to remain unmarried than females, with 30% of males unmarried compared to 26% of females.
 
@@
 
-## Insights into Education, Income, and Marriage Patterns
+## Analysis of Education, Income, and Marriage Patterns
 
@@
 
-The analysis revealed gender and racial disparities.
+The analysis shows gender and racial disparities.
 As shown in @fig-gender-marriage, men exhibit higher non-marriage rates compared to women, likely reflecting cultural expectations and differing societal pressures.
@@
 
-While this study provides valuable insights into the determinants of non-marriage, it is not without limitations.
+While this study provides useful insights into the determinants of non-marriage, it is not without limitations.
 The cross-sectional nature of the data restricts causal inferences, making it difficult to ascertain how variables such as education and income evolve over time to influence marital decisions.
@@
 
-Additionally, the dataset lacks qualitative dimensions, such as cultural attitudes or personal preferences, which are crucial for understanding the broader context of marital decisions.
+Additionally, the dataset lacks qualitative dimensions, such as cultural attitudes or personal preferences, which are necessary for understanding the broader context of marital decisions.
 For instance, while financial stability and education are strong predictors, societal expectations and individual life goals may play equally important roles that remain unmeasured.
@@
 Education emerged as the most significant predictor, with higher levels strongly associated with lower non-marriage rates.
-Income, while secondary, remains a critical factor in promoting marital stability.
+Income, while secondary, remains an essential factor in promoting marital stability.
 Racial and gender disparities further highlight systemic inequities that shape marital trends.
@@
 By addressing these structural barriers, policymakers and institutions can promote equity in family formation.
-Future research should build on these findings, leveraging longitudinal and qualitative approaches to provide a more comprehensive understanding of marriage and its evolving role in contemporary society.
+Future research should build on these findings, by employing longitudinal and qualitative approaches to provide a more extensive understanding of marriage and its evolving role in contemporary society.
 

### 2024-12-04 paper/paper.qmd
Final commit

@@
 date-format: long
-abstract: "This study examines the socio-demographic determinants of non-marriage in the United States in 2023 using logistic regression analysis on individual-level data.
-The findings highlight that individuals with lower income and education levels, particularly below a bachelor’s degree, are significantly more likely to remain unmarried.
-Additionally, gender and race also play important roles, with men and certain racial groups, such as African Americans, showing higher unmarried rates.
-These results underscore structural inequities and socio-economic barriers shaping marital outcomes, offering insights into broader societal patterns and potential areas for intervention."
+abstract: "This study examines the socio-demographic factors influencing non-marriage in the United States in 2023 through logistic regression analysis of individual-level data.
+The findings show that individuals with lower income and education levels, particularly those without a bachelor’s degree, are more likely to remain unmarried.
+Additionally, gender and race significantly influence outcomes, with men and certain racial groups, such as African Americans, experiencing higher rates of non-marriage.
+These results highlight structural inequities and socio-economic barriers affecting marital patterns and identify areas for potential policy action."
 format: pdf
@@
 The structure of this paper is as follows: @sec-data explains the dataset, variable selection, and measurement strategies employed to study marital status.
-@sec-model discusses the logistic regression methodology, including model rationale, prior specifications, and validation processes.
+@sec-model discusses the logistic regression methodology.
 @sec-results presents the key findings with supporting visualizations, emphasizing the roles of education, income, race, gender, and age in predicting non-marriage.
@@
 
-## Measurement
+## Data Source and Measurement
+
+The data employed in this research was obtained from IPUMS USA (Integrated Public Use Microdata Series) [@ipumsusa].
+This dataset contains individual-level microdata, where each record represents a distinct individual.
+The attributes of these individuals are numerically coded and span a wide array of demographic, economic, and social variables.
 
@@
 For this study,
-**Not_Married**: Includes individuals who have never been married.
-**Married**: Includes individuals who are married, as well as those who are divorced, widowed, or separated.
+
+1. **Not_Married**: Includes individuals who have never been married.
+2. **Married**: Includes individuals who are married, as well as those who are divorced, widowed, or separated.
 
@@
 
-## Determinants of Non-Marriage and Societal Implications
+## Determinants of Non-Marriage
 
-This study investigates the socio-demographic factors influencing marital status, focusing on the probability of remaining unmarried in the United States.
-Through logistic regression analysis applied to individual-level microdata, significant predictors such as education, income, race, gender, and age were identified.
+This study identifies the socio-demographic factors influencing marital status in the United States, focusing on the likelihood of remaining unmarried.
+Through logistic regression analysis, the roles of education, income, race, gender, and age were examined.
 
-Higher education levels were strongly associated with lower non-marriage rates.
-For instance, @fig-edu-marriage shows that only 16% of individuals with education above a bachelor’s degree were unmarried, compared to 32% for those with "Some College" education.
-Logistic regression coefficients further underscore this trend, with individuals holding higher degrees being substantially less likely to remain unmarried (e.g., the coefficient for "Below High School" = 0.48).
-Education thus acts as a stabilizing factor, equipping individuals with economic resources and social networks conducive to marriage.
+### The Role of Education
 
-Similarly, income plays a complementary but less pronounced role.
-@fig-Income-Unmarried illustrates that unmarried individuals tend to cluster in lower income brackets, with the majority earning less than $40,000 annually.
-This highlights the persistent importance of financial stability in marital decisions, although its impact appears secondary to education.
+Education was found to be the most significant factor associated with marital status.
+Individuals with higher educational attainment, such as those holding graduate degrees, had the lowest rates of non-marriage (16%), while those with “Some College” education exhibited a 32% likelihood of being unmarried.
+The analysis highlights how education contributes to economic stability and social networks, creating conditions that favor marriage.
+Aughinbaugh further observed that individuals with higher education are less likely to divorce and more likely to maintain stable marriages [@citeMarriagePatterns].
 
-The results also emphasize significant disparities across racial and gender lines.
-African Americans exhibited the highest non-marriage rates (43%), as shown in @fig-race-marriage, significantly surpassing those of Asian Americans (30%) and White individuals (24%).
-These patterns likely reflect structural inequalities, such as higher unemployment rates and barriers to wealth accumulation among African Americans.
-Gender differences were also evident, with men showing a slightly higher likelihood of remaining unmarried compared to women (30% versus 26%, respectively, @fig-gender-marriage).
-This aligns with cultural norms where women are more closely tied to family roles and experience greater societal pressure to marry.
+### Income as a Supporting Factor
 
-These findings reinforce how individual socio-demographic traits intersect with societal structures to shape marital outcomes.
-They also highlight systemic inequities that warrant targeted interventions.
+Income played a secondary but meaningful role in marital decisions.
+As shown in @fig-Income-Unmarried, the majority of unmarried individuals were concentrated in lower income brackets, earning less than $40,000 annually.
+This indicates that financial stability can influence decisions about marriage.
+Stable economic conditions are likely to increase individuals’ willingness to marry and form families.
+These findings highlight the relationship between economic factors and broader marriage trends.
 
-## Analysis of Education, Income, and Marriage Patterns
+## Race and Gender Differences
 
-Education emerged as the most significant predictor in this analysis.
-Higher education levels correlate strongly with lower non-marriage rates, as depicted in @fig-edu-marriage and @fig-coefficient.
-These findings align with theories suggesting that education enhances economic stability, expands social capital, and provides greater agency in life choices, all of which promote marriage.
-However, as @fig-Income-Unmarried indicates, financial stability remains an essential but secondary factor in marital decisions.
+### Racial Disparities
 
-The interplay between education and income underscores broader socio-economic disparities.
-For example, individuals with lower education levels not only face limited economic opportunities but are also more likely to remain unmarried, perpetuating cycles of inequality in family formation.
+Racial differences were pronounced in the analysis.
+African Americans had the highest non-marriage rates (43%), compared to 30% for Asian Americans and 24% for White individuals.
+These disparities may reflect systemic inequities, such as limited access to wealth and higher unemployment rates [@citeAltonjiMarriage].
+In contrast, cultural values that emphasize family cohesion may contribute to the lower rates among Asian Americans.
 
-## Gender and Race in Marital Trends
+### Gender Disparities
 
-The analysis shows gender and racial disparities.
-As shown in @fig-gender-marriage, men exhibit higher non-marriage rates compared to women, likely reflecting cultural expectations and differing societal pressures.
-Racial disparities, however, are even more striking.
-African Americans, for instance, show the highest rates of non-marriage, suggesting systemic inequities such as income disparities and historical barriers to wealth accumulation.
-By contrast, Asian Americans, with the lowest non-marriage rates, may benefit from cultural norms emphasizing family cohesion and marriage stability.
+Gender also shaped marital patterns, with men showing a slightly higher likelihood of being unmarried than women (30% vs.
+26%).
+This trend aligns with cultural expectations, where women often face greater societal pressure to marry and assume family roles [@citeMarriagePatterns].
+These findings suggest that cultural norms significantly affect marital outcomes.
 
-These patterns illustrate how cultural norms and structural inequities converge to shape marital outcomes.
-Addressing these disparities requires policies that promote economic equity and social inclusivity.
+## The Role of Age
+
+Age patterns in marital status were evident in @fig-Age-Unmarried, with younger individuals, particularly those in their 20s and early 30s, being more likely to remain unmarried.
+This likelihood decreased with age, as individuals tend to enter marriage at more mature life stages when personal and financial readiness align.
+This trend is consistent with previous research, which shows that delayed marriage has become a common demographic pattern in modern society [@citeMarriagePatterns].
 
@@
 
-While this study provides useful insights into the determinants of non-marriage, it is not without limitations.
+While this study provides detailed analysis of the determinants of non-marriage, it is not without limitations.
 The cross-sectional nature of the data restricts causal inferences, making it difficult to ascertain how variables such as education and income evolve over time to influence marital decisions.
